### 2017-11-18 paper/paper.tex
Paper turned into LaTeX
Added random references to make everything compile.

\documentclass[conference]{IEEEtran}
\usepackage{cite}
% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

\begin{document}

\title{Distributed Tic Tac Toe:\\ Peer-to-Peer Team-Based Gaming System}

\author{Panagiotis Antoniou, Emily Band, Paulius Dilkas, Dafin Kozarev, Joshua Styles}
\maketitle

\begin{abstract}
This article describes a Distributed Multiplayer Tic Tac Toe software game written in Java using the RMI API. While previous implementations attempting to serve a similar purpose exist, none of them boast good design or efficiency and neither of them deliver the same functionality. The developed piece of software allows clients to play locally or globally a game of the classic tic tac toe - but with a twist. Players are separated into teams and decide on what the next move should be by voting. The implementation is derived from a design that follows the distributed applications guidelines and is prepared to handle corner cases such as client disconnects for example.
\end{abstract}

\section{Introduction}
Coding nowadays is the main occupation of loads of people and it is also a
past-time for just as many. Most people get their first steps in writing code
following the local single-threaded model. Often, it can be difficult to
beginner programmers or even those with more experience to abstract the key
knowledge they have obtained using this model and apply it in a different way.
Distributed software requires a new look on the system – the number of devices
has gone up and with it so has the number of processes and other concerns. Some
concepts have now changed and need to be dealt with differently – one such
example is mutual exclusion. We believe programs that serve as examples and
focus on simple local tasks turned into distributed ones can help many people
trying to make this step with a hands-on experience of what they are preparing
for. Seeing theory in practice is a proven technique and there is evidence for
this in every university course that features practical work. Therefore, we have
chosen to build a distributed game of tic-tac-toe – a game known to possibly
everyone. By restraining from introducing additional complexity as part of the
goal of the project, the focus is shifted towards the implications of the
distributive aspect of the system and the way they are handled. The finished
product also has another purpose – it is an actual game that can be played and
as such it can bring joy to people and increase their curiosity about computers.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{System Design}

\section{Implementation}

\section{Evaluation}

\section{Related Work}
While other multiplayer distributed games already exist, they have not been
created with this project’s task in mind and therefore do not serve the same
purpose. They can be split into two categories – very unprofessional and too
professional. For example, hobby implementations exist that deal with
distributed problems in a bad fashion. Similarly, a single-player-per-team
distributed tic-tac-toe game in Java can be found online that does not comply
with any design guidelines and puts all functionality in a single class and as
such cannot teach students good practice and might confuse them further about
distributive programming \cite{2}. Separately, very in-depth materials that
provide state-of-the-art practices can be found \cite{1} but there is a high
possibility that they might be incomprehensible to a beginner reader looking to
tackle basic distributed concepts to begin with. Furthermore, none of the
similar implementations residing on the Internet take advantage of the Java RMI
API which must be considered a good starting point for distributed systems
learners for it to be taught as part of our degree and is a requirement for this
project. Instead, they use different sets of other technologies which can
further confuse their reader and introduce a steeper learning curve.

\section{Conclusions and Future Work}

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
\bibliography{references}
\end{document}

### 2017-11-18 Game.java
Adding Dafin's changes with some modifications

@@ -2,26 +2,29 @@ import java.io.IOException;
 import java.rmi.Naming;
 import java.rmi.NotBoundException;
 import java.rmi.RemoteException;
-import java.rmi.registry.LocateRegistry;
-import java.rmi.registry.Registry;
 import java.rmi.server.UnicastRemoteObject;
 import java.util.LinkedList;
 import java.util.List;
 import java.util.Scanner;
 
 public class Game extends UnicastRemoteObject implements GameInt {
-    // TODO: Reorder the methods. Right now the order is chronological/arbitrary.
-    // TODO: Make role public
-    // TODO: More (and more accurate) comments
 
+    /** Serial version UID to avoid warning. */
+    private static final long serialVersionUID = 42L;
+    /** Address of host machine. */
     private static final String HOST = "localhost";
+    /** Port number of the game. */
     private static final int PORT = 1099;
 
+    /** Holds the combined IP, port and server/client name address of this instance. */
     private String myUrl;
     private Scanner input;
     private GameState gameState;
+    /** Identifies the role of this instance. */
     private Role role;
 
+
+    /** Constructors */
     public Game(String url) throws RemoteException {
         myUrl = url;
         input = new Scanner(System.in);
@@ -35,13 +38,8 @@ public class Game extends UnicastRemoteObject implements GameInt {
         this.gameState = gameState;
     }
 
-    public boolean addPlayer(GameInt player) throws RemoteException {
-        return role.addPlayer(player);
-    }
-
-    /** Asks the user to make a play and returns it. */
-    public int yourTurn() throws RemoteException {
-        // TODO: perhaps rename this function to something more appropriate
+    /** Ask the user to make a play and return it. */
+    public int askForInput() throws RemoteException {
         int play = -1;
         while (!gameState.valid(play)) {
             System.out.print("Your sign: " + gameState.nextSign + ". Choose a number in [0, 8]: ");
@@ -50,43 +48,38 @@ public class Game extends UnicastRemoteObject implements GameInt {
         return play;
     }
 
-    /** Receives a play and updates the game state. Returns true if the game is over. */
+    /** Receive a play and update the game state. Return true if the game is over. */
     public boolean makePlay(int play) throws RemoteException {
         boolean gameOver = gameState.makePlay(play);
         System.out.println(gameState);
         return gameOver;
     }
 
+    /** End the game. */
     public void endGame() throws RemoteException {
         input.close();
         try {
             Naming.unbind(myUrl);
[…]

### 2017-11-18 GameInt.java

@@ -5,7 +5,7 @@ public interface GameInt extends Remote {
 
     public boolean addPlayer(GameInt player) throws RemoteException;
 
-    public int yourTurn() throws RemoteException;
+    public int askForInput() throws RemoteException;
 
     public boolean makePlay(int play) throws RemoteException;
 

### 2017-11-18 GameState.java

@@ -2,6 +2,9 @@ import java.io.Serializable;
 
 public class GameState implements Serializable {
 
+    /** Serial version UID to avoid warning. */
+    private static final long serialVersionUID = 53L;
+
     private enum Sign {
         EMPTY, X, O;
     };

### 2017-11-18 Leader.java

@@ -1,13 +1,14 @@
 import java.rmi.RemoteException;
 import java.util.ArrayList;
-import java.util.LinkedList;
 import java.util.List;
 import java.util.Random;
+import java.util.concurrent.atomic.AtomicIntegerArray;
 
 public class Leader extends Role {
 
     // The local Game instance is here as well.
     // Synchronize whenever using this list.
+    // TODO: use a data structure that support concurrency and remove all the 'synchronized' keywords
     private List<GameInt> myTeam;
 
     public Leader(List<GameInt> team) {
@@ -23,24 +24,38 @@ public class Leader extends Role {
         return false;
     }
 
+    /** Collect the votes from the team and decide on a play. */
     public void turnStarts() throws RemoteException {
-        // Gather the votes
-        int[] votes = new int[9];
+        AtomicIntegerArray votes = new AtomicIntegerArray(9);
+        int counter = 0;
+        VoteThread[] voteCollectors;
         synchronized (this) {
-            // TODO: make this asynchronous (important)
+            int teamSize = myTeam.size();
+            voteCollectors = new VoteThread[teamSize];
             for (GameInt player : myTeam)
-                votes[player.yourTurn()]++;
+                voteCollectors[counter++] = new VoteThread(player, votes);
+        }
+
+        for (VoteThread thread : voteCollectors)
+            thread.start();
+
+        for (VoteThread thread : voteCollectors) {
+            try {
+                thread.join();
+            } catch (InterruptedException e) {
+                e.printStackTrace();
+            }
         }
 
         // Choose the winning play, breaking ties randomly
         List<Integer> maxIndices = new ArrayList<>();
         int maxVotes = 0;
         for (int i = 0; i < 9; i++) {
-            if (votes[i] > maxVotes) {
+            if (votes.get(i) > maxVotes) {
                 maxIndices = new ArrayList<>();
-                maxVotes = votes[i];
+                maxVotes = votes.get(i);
             }
-            if (votes[i] == maxVotes)
+            if (votes.get(i) == maxVotes)
                 maxIndices.add(i);
         }
         int play = maxIndices.get(new Random().nextInt(maxIndices.size()));

### 2017-11-18 VoteThread.java

@@ -1,0 +1,21 @@
+import java.rmi.RemoteException;
+import java.util.concurrent.atomic.AtomicIntegerArray;
+
+/** Gets a vote from a player. */
+class VoteThread extends Thread {
+    private GameInt player;
+    private AtomicIntegerArray votes;
+
+    public VoteThread(GameInt player, AtomicIntegerArray votes) {
+        this.player = player;
+        this.votes = votes;
+    }
+
+    public void run() {
+        try {
+            votes.getAndIncrement(player.askForInput());
+        } catch (RemoteException e) {
+            e.printStackTrace();
+        }
+    }
+}

### 2017-11-19 paper/paper.tex
Article Additions
"Article Guide" document to help fill out each section, as per the assignment spec.

"References" document to keep track of references along with citation number until we can index it ourselves.

Small changes to the paper to make it less informal. Still more needed on this.

Small additions to Introduction to argue "novel contribution" and high level overview of the game in the "System Design" section.

@@
 \section{Introduction}
-Coding nowadays is the main occupation of loads of people and it is also a
-past-time for just as many. Most people get their first steps in writing code
-following the local single-threaded model. Often, it can be difficult to
-beginner programmers or even those with more experience to abstract the key
-knowledge they have obtained using this model and apply it in a different way.
+Coding is the main occupation of many people nowadays, and is also a
+past-time for many too. Most people get their first steps in writing code
+following the local single-threaded model. Often, it can be difficult for
+beginner programmers, or even those with more experience, to abstract the key
+knowledge they have obtained using this model and then apply it in a different way.
+
 Distributed software requires a new look on the system – the number of devices
-has gone up and with it so has the number of processes and other concerns. Some
-concepts have now changed and need to be dealt with differently – one such
-example is mutual exclusion. We believe programs that serve as examples and
+has gone up and with it so have the number of processes, concerns and complexities.
+Some concepts have now changed and need to be dealt with differently; One such
+example if this would be mutual exclusion. We believe programs that serve as examples and
 focus on simple local tasks turned into distributed ones can help many people
@@
 for. Seeing theory in practice is a proven technique and there is evidence for
-this in every university course that features practical work. Therefore, we have
-chosen to build a distributed game of tic-tac-toe – a game known to possibly
-everyone. By restraining from introducing additional complexity as part of the
-goal of the project, the focus is shifted towards the implications of the
-distributive aspect of the system and the way they are handled. The finished
-product also has another purpose – it is an actual game that can be played and
-as such it can bring joy to people and increase their curiosity about computers.
+this in every university course that features practical work. This is especially
+true where games are concerned, proving to be an effective means of communicating 
+programming methods and techniques in an engaging way \cite{3}. For this reason, 
+the team has decided to take a simple, well known game and develop it into a distributed 
+system; Tic Tac Toe played as a peer-to-peer, team-based gaming system where the players are organised into teams and required to vote on moves for their team.
+
+By refraining from introducing additional complexity as part of the goal of the project, and 
+allowing the users to see the distributed nature of the software being demonstrated via a simple game, the focus is shifted towards the implications of the distributive aspects of the system and 
+the way they are handled.
+The finished product can serve to entertain the end users, as well as increase their curiosity and insight about computers and distributed systems.
+Using Tic Tac Toe as a foundation to demonstrate a distributed system benefits from both the simplicity and familiarity of the game itself, meaning no complex rulesets or in-depth mechanisms need be explained to the users, allowing the distributed nature to be a novel way of experiencing a widely known game.
 
@@
 \section{System Design}
+Tic Tac Toe is a simple game played on a grid of 3x3, consisting of two players.
+One player uses an “X” marker and the other player uses an “O” marker to select a grid space to occupy.
+The objective of the game is to take turns until one player manages to occupy three grid spaces in a straight line, be it horizontal, vertical or diagonal.
+
+The proposed design takes this simple concept and uses it to demonstrate a distributed system, whereby the two “players” are instead two teams of up to 5 players, which take turns to vote on where their team should place their markers.
+Each team has a designated "leader" which communicates data with their team's peers, and the opposing team's "leader" to synchronise current gameplay state.
+The leaders communicate the vote of their team to the opposing leader so that they may pass that result on to the peers of that team.
 
@@
 professional. For example, hobby implementations exist that deal with
-distributed problems in a bad fashion. Similarly, a single-player-per-team
-distributed tic-tac-toe game in Java can be found online that does not comply
-with any design guidelines and puts all functionality in a single class and as
-such cannot teach students good practice and might confuse them further about
-distributive programming \cite{2}. Separately, very in-depth materials that
-provide state-of-the-art practices can be found \cite{1} but there is a high
-possibility that they might be incomprehensible to a beginner reader looking to
-tackle basic distributed concepts to begin with. Furthermore, none of the
-similar implementations residing on the Internet take advantage of the Java RMI
-API which must be considered a good starting point for distributed systems
+distributed problems in a bad fashion, promoting bad and incorrect programming 
+practises. Similarly, a single-player-per-team distributed tic-tac-toe game in 
+Java can be found online that does not comply with any design guidelines and 
+puts all functionality in a single class and as such, cannot teach students good 
+practice and has the potential of confusing the learner further about distributive 
+programming \cite{2}.
+
+Separately, very in-depth materials that provide state-of-the-art practices can 
+be found \cite{1} but there is a high possibility that they might be incomprehensible 
+to a beginner reader looking to tackle basic distributed concepts to begin with. 
+Furthermore, none of the similar implementations residing on the Internet take advantage 
+of the Java RMI API which must be considered a good starting point for distributed systems
 learners for it to be taught as part of our degree and is a requirement for this

### 2017-11-19 BullyAlgorithmInt.java
Updated bully algoritm

@@ -1,18 +1,0 @@
-import java.rmi.Remote;
-import java.rmi.RemoteException;
-
-public interface BullyAlgorithmInt extends Remote{
-
-    // Leaders of each team will be the oldest players joined the game
-    // If player not oldest:
-    //       if timeoutElection then consider itself coordinator
-    //       if timeoutCoordinator then start another election
-    public int timeoutElection = 10, timeoutCoordinator = 10;
-
-    public void startElection() throws RemoteException;
-
-    public void responseToElection() throws RemoteException;
-
-    public void announceCoordinator() throws RemoteException;
-
-}

### 2017-11-19 GameInt.java

@@ -13,11 +13,37 @@ public interface GameInt extends Remote {
 
     public GameState getGameState() throws RemoteException;
 
-    public void setLeader() throws RemoteException;
+//    public void setLeader() throws RemoteException;
+
+//    public void setLeader(GameInt leader,GameInt opponent) throws RemoteException;
 
     public void turnStarts() throws RemoteException;
 
     public void printBoard() throws RemoteException;
 
     public void endGame() throws RemoteException;
+
+
+
+
+    /**
+     *
+     *  Bully Algorithm Interface
+     *
+     * */
+
+    // Leaders of each team will be the oldest players joined the game
+    // If player not oldest:
+    //       if timeoutElection then consider itself coordinator
+    //       if timeoutCoordinator then start another election
+    public int timeoutElection = 10, timeoutCoordinator = 10;
+
+
+    // Use this to start an election from oldest processes than you
+    // Response is used for the timeout in the calling process
+    public boolean startElection() throws RemoteException;
+
+
+    // Use this to broadcast to all the processes in your team & opponent leader , about the new leader
+    public void declareLeader(GameInt leader) throws RemoteException;
 }

### 2017-11-19 Leader.java

@@ -20,7 +20,7 @@ public class Leader extends Role {
             leader = player;
             return true;
         }
-        team.add(player);
+//        team.add(player);
         return false;
     }
 

### 2017-11-19 Role.java

@@ -1,8 +1,12 @@
 import java.rmi.RemoteException;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.Timer;
+import java.util.TimerTask;
 import java.util.concurrent.PriorityBlockingQueue;
 
 /** State design pattern */
-public abstract class Role {
+public abstract class Role{
 
     protected GameInt leader;                          // Current Leader
     protected GameInt opponentLeader;                  // Current Opponent Leader
@@ -15,18 +19,54 @@ public abstract class Role {
         this.team = team;
     }
 
-    /** Returns true if the added player is a leader */
-    public abstract boolean addPlayer(GameInt player) throws RemoteException;
-
-    /** A leader overrides this, a regular player does nothing */
-    public void broadcastPlay(int play) throws RemoteException {
-    }
 
     public void setLeader(GameInt somePlayer) {
         leader = somePlayer;
     }
 
+    /** Returns true if the added player is a leader */
+    public abstract boolean addPlayer(GameInt player) throws RemoteException;
+
+    /** Returns true if the added player is an opponent leader*/
+    public synchronized boolean addOpponent(GameInt player) throws RemoteException{
+        if (opponentLeader == null) {
+            opponentLeader = player;
+            return true;
+        }
+        team.add(player);
+        return false;
+    }
+
+    /** A leader overrides this, a regular player does nothing */
+    public void broadcastPlay(int play) throws RemoteException {
+    }
+
     /** Regular players ignore this. */
     public void turnStarts() throws RemoteException {
+    }
+
+    // False indicates leader has no "greater" leaders
+    // True indicates leader has started another election
+    // No response assume other process crashed -- Use timeout when calling it
+    public boolean startElection() throws RemoteException {
+        // Start an election to my team
+        GameInt[] myTeam = (GameInt[]) this.team.toArray();
+        ArrayList<Boolean> responses = new ArrayList<>();
+        int currentGameInt = 0;
+
+        while(this.hashCode() < myTeam[currentGameInt].hashCode() && (currentGameInt<myTeam.length)){
+            responses.add(myTeam[currentGameInt].startElection());
+            currentGameInt++;
+        }
+        // No more leaders set me as leader
+        if (responses.size() == 0){
+            return false;
+        }
+        return true;
+    }
+
+
+    public void declareLeader(GameInt leader) throws RemoteException {
+        this.leader = leader;
     }
 }

### 2017-11-19 Game.java
Updated Game.java from master and added new methods for bully algorithm, modified rest of the files accordingly

@@ -10,7 +10,6 @@ import java.util.List;
 import java.util.Scanner;
 import java.util.concurrent.PriorityBlockingQueue;
 
-
 public class Game extends UnicastRemoteObject implements GameInt {
     // TODO: Reorder the methods. Right now the order is chronological/arbitrary.
     // TODO: Make role public
@@ -23,27 +22,19 @@ public class Game extends UnicastRemoteObject implements GameInt {
     private Scanner input;
     private GameState gameState;
     private Role role;
-//    private int id;
-//    private static int idCount = 0;
 
     public Game(String url) throws RemoteException {
         myUrl = url;
         input = new Scanner(System.in);
         gameState = new GameState();
-//        id = idCount++;
         setLeader();
     }
 
     public Game(String url, GameState gameState) throws RemoteException {
         myUrl = url;
         input = new Scanner(System.in);
-//        id = idCount++;
         this.gameState = gameState;
     }
-
-//    public int getId(){
-//        return id;
-//    }
 
     public boolean addPlayer(GameInt player) throws RemoteException {
         return role.addPlayer(player);
@@ -79,6 +70,16 @@ public class Game extends UnicastRemoteObject implements GameInt {
         System.out.println("Game over.");
     }
 
+    @Override
+    public boolean startElection() throws RemoteException {
+        return role.startElection();
+    }
+
+    @Override
+    public void declareLeader(GameInt leader) throws RemoteException {
+        role.leader = leader;
+    }
+
     public void broadcastPlay(int play) throws RemoteException {
         role.broadcastPlay(play);
     }
@@ -88,15 +89,15 @@ public class Game extends UnicastRemoteObject implements GameInt {
     }
 
     /** Make this instance into a leader and initialise the team list */
-    public void setLeader(GameInt leader,GameInt opponentLeader,PriorityBlockingQueue<GameInt> team) throws RemoteException {
-        this.role = new Leader(leader,opponentLeader,team);
+    public void setLeader(PriorityBlockingQueue<GameInt> team) throws RemoteException {
+        this.role = new Leader(team);
         System.out.println("I am a leader now.");
     }
 
     public void setLeader() throws RemoteException {
-        PriorityBlockingQueue<GameInt> initialTeam = new PriorityBlockingQueue<>(10,new GameIntComparator());
+        PriorityBlockingQueue<GameInt> initialTeam = new PriorityBlockingQueue<>();
         initialTeam.add(this);
-        setLeader(this,null,initialTeam);
[…]

### 2017-11-19 GameInt.java

@@ -13,7 +13,7 @@ public interface GameInt extends Remote {
 
     public GameState getGameState() throws RemoteException;
 
-//    public void setLeader() throws RemoteException;
+    public void setLeader() throws RemoteException;
 
 //    public void setLeader(GameInt leader,GameInt opponent) throws RemoteException;
 

### 2017-11-19 Leader.java

@@ -11,6 +11,10 @@ public class Leader extends Role {
     // Synchronize whenever using this list.
     // private List<GameInt> myTeam; --> moved to Role
 
+    public Leader(PriorityBlockingQueue<GameInt> team) {
+        this.team = team;
+    }
+
     public Leader(GameInt leader,GameInt opponentLeader,PriorityBlockingQueue<GameInt> team) {
         super(leader,opponentLeader,team);
     }

### 2017-11-19 Role.java

@@ -12,6 +12,11 @@ public abstract class Role{
     protected GameInt opponentLeader;                  // Current Opponent Leader
     protected PriorityBlockingQueue<GameInt> team;     // My Team
 
+    // Temporary for conflicting Leader call constructor
+    public Role(PriorityBlockingQueue<GameInt> team){
+        this.team = team;
+    }
+
     /** Used by the subclasses */
     public Role(GameInt leader,GameInt opponentLeader,PriorityBlockingQueue<GameInt> team){
         this.leader = leader;

### 2017-11-20 paper/paper.tex
Article Update
Article now includes diagram of current topology as provided by Emily.

@@
 % Basically, \url{my_url_here}.
+\usepackage{graphicx}
 
@@
 Each team has a designated "leader" which communicates data with their team's peers, and the opposing team's "leader" to synchronise current gameplay state.
-The leaders communicate the vote of their team to the opposing leader so that they may pass that result on to the peers of that team.
 
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{images/DAS-topology.png}
+	\caption{Topology for Distributed Tic Tac Toe}
+	\label{fig:topology1}
+\end{figure}
+
+The leaders communicate the vote of their team to the opposing leader so that they may pass that result on to the peers of that team, as pictured in figure 1.
 \section{Implementation}

### 2017-11-20 paper/paper.tex
Fixed a small bug and added a reference

@@
 true where games are concerned, proving to be an effective means of communicating 
-programming methods and techniques in an engaging way \cite{3}. For this reason, 
-the team has decided to take a simple, well known game and develop it into a distributed 
-system; Tic Tac Toe played as a peer-to-peer, team-based gaming system where the players are organised into teams and required to vote on moves for their team.
+programming methods and techniques in an engaging way
+\cite{Kiniry:2011:VG:1984674.1984681}. For this reason, the team has decided to
+take a simple, well known game and develop it into a distributed system; Tic Tac
+Toe played as a peer-to-peer, team-based gaming system where the players are
+organised into teams and required to vote on moves for their team.
 

### 2017-11-20 Game.java

@@ -36,6 +36,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
         myUrl = url;
         input = new Scanner(System.in);
         this.gameState = gameState;
+        role = new Player();
     }
 
     /** Ask the user to make a play and return it. */

### 2017-11-21 Game.java
added Timer, Paper - Implementation, References; changed addPlayer
Regular players can now detect if a leader has timed out and begin election (currently commented out as election still unimplemented). addPlayer() now returns a GameInt object not a boolean and a Player is correctly set as a role if a Leader is present. This was needed as the solution added during the last meeting did not work. Some content was added to the Implementation section of the paper. Two references were added, one is from Youtube, I hope it works with the current syntax.

@@ -36,12 +36,14 @@ public class Game extends UnicastRemoteObject implements GameInt {
         myUrl = url;
         input = new Scanner(System.in);
         this.gameState = gameState;
-        role = new Player();
     }
 
     /** Ask the user to make a play and return it. */
     public int askForInput() throws RemoteException {
-        int play = -1;
+        if(role.getClass().getName() == "Player") {
+        	role.schedule();
+        } 
+    	int play = -1;
         while (!gameState.valid(play)) {
             System.out.print("Your sign: " + gameState.nextSign + ". Choose a number in [0, 8]: ");
             play = input.nextInt();
@@ -51,7 +53,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
 
     /** Receive a play and update the game state. Return true if the game is over. */
     public boolean makePlay(int play) throws RemoteException {
-        boolean gameOver = gameState.makePlay(play);
+    	boolean gameOver = gameState.makePlay(play);
         System.out.println(gameState);
         return gameOver;
     }
@@ -79,7 +81,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
     }
 
     /** Role access classes */
-    public boolean addPlayer(GameInt player) throws RemoteException {
+    public GameInt addPlayer(GameInt player) throws RemoteException {
         return role.addPlayer(player);
     }
 
@@ -95,17 +97,22 @@ public class Game extends UnicastRemoteObject implements GameInt {
         role.broadcastPlay(play);
     }
 
-    /** Make this instance into a leader and initialise the team list. */
+    /** Make this instance into a leader of a brand new game. */
     public void setLeader(List<GameInt> team) throws RemoteException {
         this.role = new Leader(team);
         System.out.println("I am a leader now.");
     }
 
-    /** Make this instance into a leader of a brand new game. */
+    /** Make this instance into a leader and initialise the team list. */
     public void setLeader() throws RemoteException {
         List<GameInt> initialTeam = new LinkedList<>();
         initialTeam.add(this);
         setLeader(initialTeam);
+    }
+    
+    /** Make this instance into a regular player */
+    public void setAsPlayer(GameInt leader) throws RemoteException {
+        this.role = new Player(leader);
     }
 
     public static void main(String[] args)throws IOException {
@@ -125,13 +132,15 @@ public class Game extends UnicastRemoteObject implements GameInt {
             GameInt server = (GameInt) Naming.lookup("//" + args[2] + ":" + PORT + "/" + args[1]);
[…]

### 2017-11-21 GameInt.java

@@ -3,7 +3,7 @@ import java.rmi.RemoteException;
 
 public interface GameInt extends Remote {
 
-    public boolean addPlayer(GameInt player) throws RemoteException;
+    public GameInt addPlayer(GameInt player) throws RemoteException;
 
     public int askForInput() throws RemoteException;
 
@@ -14,6 +14,8 @@ public interface GameInt extends Remote {
     public GameState getGameState() throws RemoteException;
 
     public void setLeader() throws RemoteException;
+    
+    public void setAsPlayer(GameInt leader) throws RemoteException;
 
     public void turnStarts() throws RemoteException;
 

### 2017-11-21 Leader.java

@@ -7,7 +7,7 @@ import java.util.concurrent.atomic.AtomicIntegerArray;
 public class Leader extends Role {
 
     // The local Game instance is here as well.
-    // Synchronize whenever using this list.
+    // Synchronise whenever using this list.
     // TODO: use a data structure that support concurrency and remove all the 'synchronized' keywords
     private List<GameInt> myTeam;
 
@@ -15,13 +15,13 @@ public class Leader extends Role {
         myTeam = team;
     }
 
-    public synchronized boolean addPlayer(GameInt player) throws RemoteException {
+    public synchronized GameInt addPlayer(GameInt player) throws RemoteException {
         if (leader == null) {
             leader = player;
-            return true;
+            return null;
         }
         myTeam.add(player);
-        return false;
+        return (GameInt) this;
     }
 
     /** Collect the votes from the team and decide on a play. */
@@ -88,4 +88,9 @@ public class Leader extends Role {
             myTeam = null;
         }
     }
+
+	public void schedule() {
+		// Schedule Timer to record last response of each regular player on team
+		// If idle for more than x minutes, delete from team and inform other members to remove reference
+	}
 }

### 2017-11-21 Player.java

@@ -1,13 +1,49 @@
 import java.rmi.RemoteException;
+import java.util.Calendar;
+import java.util.Random;
+import java.util.Timer;
+import java.util.TimerTask;
 
 public class Player extends Role {
-
+	private Timer timer;
+	
     public Player(GameInt leader) {
         this.leader = leader;
     }
 
-    public boolean addPlayer(GameInt player) throws RemoteException {
+    public GameInt addPlayer(GameInt player) throws RemoteException {
         leader.addPlayer(player);
-        return false;
+        return leader;
+    }
+    
+    public void schedule() {
+    	timer.cancel();
+    	long delay, period;
+        Calendar cal=Calendar.getInstance(); // get today
+        int hr = cal.get(Calendar.HOUR_OF_DAY);
+        int mn = cal.get(Calendar.MINUTE);
+        mn = 5 * (mn / 5 + 1);	// round up to next 5 minute slot
+        if (mn == 60) {
+          hr = hr + 1; mn = 0;
+        }
+        cal.set(Calendar.HOUR_OF_DAY,hr);
+        cal.set(Calendar.MINUTE,mn);
+        cal.set(Calendar.SECOND,0);
+        // milliseconds until next 5 minute boundary
+        delay = cal.getTimeInMillis() - System.currentTimeMillis();
+        Random rand = new Random();
+        int randInt = rand.nextInt(500) + 1;
+        period = 5 * 60 * 1000 + randInt; // 5 minutes in milliseconds plus a random extension from 1 to 500ms
+        timer.schedule(new ElectionRaiser(), delay, period);
+    	timer.purge();
+    }
+    
+    /**
+     * task that is scheduled by the timer
+     */
+    class ElectionRaiser extends TimerTask {
+      public void run() {
+        //raiseElection();
+      }
     }
 }

### 2017-11-21 Role.java

@@ -8,7 +8,7 @@ public abstract class Role {
     protected GameInt leader;
 
     /** Returns true if the added player is a leader */
-    public abstract boolean addPlayer(GameInt player) throws RemoteException;
+    public abstract GameInt addPlayer(GameInt player) throws RemoteException;
 
     /** A leader overrides this, a regular player does nothing */
     public void broadcastPlay(int play) throws RemoteException {
@@ -17,6 +17,8 @@ public abstract class Role {
     public void setLeader(GameInt somePlayer) {
         leader = somePlayer;
     }
+    
+    public abstract void schedule();
 
     /** Regular players ignore this. */
     public void turnStarts() throws RemoteException {

### 2017-11-22 Game.java
Update Game.java

@@ -40,7 +40,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
 
     /** Ask the user to make a play and return it. */
     public int askForInput() throws RemoteException {
-        if(role.getClass().getName() == "Player") {
+        if(role instanceof Player) {
         	role.schedule();
         } 
     	int play = -1;

### 2017-11-22 Game.java
Some improvements in both functionality and style
Works at least well enough to merge into the main branch.

@@ -40,10 +40,9 @@ public class Game extends UnicastRemoteObject implements GameInt {
 
     /** Ask the user to make a play and return it. */
     public int askForInput() throws RemoteException {
-        if(role instanceof Player) {
-        	role.schedule();
-        } 
-    	int play = -1;
+        if(role instanceof Player)
+            role.schedule();
+        int play = -1;
         while (!gameState.valid(play)) {
             System.out.print("Your sign: " + gameState.nextSign + ". Choose a number in [0, 8]: ");
             play = input.nextInt();
@@ -53,7 +52,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
 
     /** Receive a play and update the game state. Return true if the game is over. */
     public boolean makePlay(int play) throws RemoteException {
-    	boolean gameOver = gameState.makePlay(play);
+        boolean gameOver = gameState.makePlay(play);
         System.out.println(gameState);
         return gameOver;
     }
@@ -85,10 +84,6 @@ public class Game extends UnicastRemoteObject implements GameInt {
         return role.addPlayer(player);
     }
 
-    public void setLeader(GameInt somePlayer) throws RemoteException {
-        role.setLeader(somePlayer);
-    }
-
     public void turnStarts() throws RemoteException {
         role.turnStarts();
     }
@@ -97,7 +92,12 @@ public class Game extends UnicastRemoteObject implements GameInt {
         role.broadcastPlay(play);
     }
 
-    /** Make this instance into a leader of a brand new game. */
+    /** Update the leader reference. */
+    public void setLeader(GameInt somePlayer) throws RemoteException {
+        role.setLeader(somePlayer);
+    }
+
+    /** Make this instance into a leader with a given team. */
     public void setLeader(List<GameInt> team) throws RemoteException {
         this.role = new Leader(team);
         System.out.println("I am a leader now.");
@@ -109,8 +109,8 @@ public class Game extends UnicastRemoteObject implements GameInt {
         initialTeam.add(this);
         setLeader(initialTeam);
     }
-    
-    /** Make this instance into a regular player */
+
+    /** Make this instance into a regular player and initialise the leader reference. */
     public void setAsPlayer(GameInt leader) throws RemoteException {
         this.role = new Player(leader);
     }
@@ -140,7 +140,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
                 game.printBoard();
                 game.turnStarts();
             } else {
-            	game.setAsPlayer(leader);
[…]

### 2017-11-22 GameInt.java

@@ -14,7 +14,7 @@ public interface GameInt extends Remote {
     public GameState getGameState() throws RemoteException;
 
     public void setLeader() throws RemoteException;
-    
+
     public void setAsPlayer(GameInt leader) throws RemoteException;
 
     public void turnStarts() throws RemoteException;

### 2017-11-22 Leader.java

@@ -15,13 +15,14 @@ public class Leader extends Role {
         myTeam = team;
     }
 
+    /** Adds a new player to the game. Returns null if the new player becomes the enemy leader and the  */
     public synchronized GameInt addPlayer(GameInt player) throws RemoteException {
         if (leader == null) {
             leader = player;
             return null;
         }
         myTeam.add(player);
-        return (GameInt) this;
+        return myTeam.get(0);
     }
 
     /** Collect the votes from the team and decide on a play. */

### 2017-11-22 Player.java

@@ -5,45 +5,49 @@ import java.util.Timer;
 import java.util.TimerTask;
 
 public class Player extends Role {
-	private Timer timer;
-	
+
+    private Timer timer;
+    private TimerTask task;
+
     public Player(GameInt leader) {
         this.leader = leader;
+        timer = new Timer();
     }
 
     public GameInt addPlayer(GameInt player) throws RemoteException {
         leader.addPlayer(player);
         return leader;
     }
-    
+
     public void schedule() {
-    	timer.cancel();
-    	long delay, period;
+        if (task != null)
+            task.cancel();
+
+        long delay, period;
         Calendar cal=Calendar.getInstance(); // get today
         int hr = cal.get(Calendar.HOUR_OF_DAY);
         int mn = cal.get(Calendar.MINUTE);
         mn = 5 * (mn / 5 + 1);	// round up to next 5 minute slot
         if (mn == 60) {
-          hr = hr + 1; mn = 0;
+            hr = hr + 1;
+            mn = 0;
         }
-        cal.set(Calendar.HOUR_OF_DAY,hr);
-        cal.set(Calendar.MINUTE,mn);
-        cal.set(Calendar.SECOND,0);
+        cal.set(Calendar.HOUR_OF_DAY, hr);
+        cal.set(Calendar.MINUTE, mn);
+        cal.set(Calendar.SECOND, 0);
         // milliseconds until next 5 minute boundary
         delay = cal.getTimeInMillis() - System.currentTimeMillis();
         Random rand = new Random();
         int randInt = rand.nextInt(500) + 1;
         period = 5 * 60 * 1000 + randInt; // 5 minutes in milliseconds plus a random extension from 1 to 500ms
-        timer.schedule(new ElectionRaiser(), delay, period);
-    	timer.purge();
+        task = new ElectionRaiser();
+        timer.schedule(task, delay, period);
     }
-    
-    /**
-     * task that is scheduled by the timer
-     */
-    class ElectionRaiser extends TimerTask {
-      public void run() {
-        //raiseElection();
-      }
+
+    /** Task that is scheduled by the timer */
+    private class ElectionRaiser extends TimerTask {
+        public void run() {
+            //raiseElection();
+        }
     }
 }

### 2017-11-22 Role.java

@@ -17,7 +17,7 @@ public abstract class Role {
     public void setLeader(GameInt somePlayer) {
         leader = somePlayer;
     }
-    
+
     public abstract void schedule();
 
     /** Regular players ignore this. */

### 2017-11-22 paper/paper.tex

@@
 \section{Introduction}
-Coding is the main occupation of many people nowadays, and is also a
-past-time for many too. Most people get their first steps in writing code
+Coding is the main occupation of many people nowadays, and it is a
+past-time for many too. Most beginners get their first steps in writing code
 following the local single-threaded model. Often, it can be difficult for
-beginner programmers, or even those with more experience, to abstract the key
+new programmers, or even those with more experience, to abstract the key
 knowledge they have obtained using this model and then apply it in a different way.
@@
 \section{Implementation}
-
+The implementation follows the RMI distributed object model. A GameInt interface extends java.rmi.Remote and lists remote classes. 
+A Game class implements GameInt and extends the UnicastRemoteObject. It is an analogy for what a player knows and does during a game. 
+It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, 
+setting the leader of the game, several methods to interact with the Role of the player and a main method that connects to other game instances. 
+Currently, when running the game, the user passes it a local instance name and the name of another instance as arguments. Each instance then uses this information 
+to do an RMI Registry lookup and get a reference to the other client. The Role interface allows the application to differentiate between Leader and Player classes which extend it. 
+It utilises the State design pattern to allow players to seamlessly have their role changed – for example when the current leader disconnects. 
+A Leader is responsible of collecting the votes from all of the members in their team in the turnStarts() method. This is done asynchronously using
+a separate thread for each member – an instance of VoteThread. Each thread collects a vote from one assigned player then writes it into a corresponding position
+in a shared AtomicIntegerArray that they have all been passed by the Leader. This ensures the main Leader thread can then read these values. 
+Each Leader then informs all players in the game about the vote of his team. A Player instance represents a regular player that only must vote. 
+Each regular player also has a mechanism to detect if a leader has not contacted them to choose a play for too long. If a time-out is triggered, 
+a player calls for a new leader election. To avoid many simultaneous election calls, each regular player waits 5 minutes plus a randomised amount of time.
+All plays are recorded on a GameState object instance that acts as a game board. After each turn it checks if the game is over and relays this message to the leaders
+ which can then inform the members of their team.
 \section{Evaluation}
@@
 practice and has the potential of confusing the learner further about distributive 
-programming \cite{2}.
+programming \cite{MrWayFarOutYoutube15}.
 
 Separately, very in-depth materials that provide state-of-the-art practices can 
-be found \cite{1} but there is a high possibility that they might be incomprehensible 
+be found \cite{DBLP:conf/nsdi/BharambePS06} but there is a high possibility that they might be incomprehensible 
 to a beginner reader looking to tackle basic distributed concepts to begin with. 

### 2017-11-22 Game.java
Sometimes automatic merging fails

@@ -36,7 +36,6 @@ public class Game extends UnicastRemoteObject implements GameInt {
         myUrl = url;
         input = new Scanner(System.in);
         this.gameState = gameState;
-        role = new Player();
     }
 
     /** Ask the user to make a play and return it. */

### 2017-11-22 Game.java
Ending the game seems to work!
Turns out debugging multithreaded Java code can be complicated. The
solutions is simple, though.

@@ -54,19 +54,18 @@ public class Game extends UnicastRemoteObject implements GameInt {
     public boolean makePlay(int play) throws RemoteException {
         boolean gameOver = gameState.makePlay(play);
         System.out.println(gameState);
-        return gameOver;
-    }
-
-    /** End the game. */
-    public void endGame() throws RemoteException {
-        input.close();
-        try {
-            Naming.unbind(myUrl);
-        } catch (Exception e) {
-            e.printStackTrace();
+        if (gameOver) {
+            input.close();
+            role.cancelTimer();
+            try {
+                Naming.unbind(myUrl);
+            } catch (Exception e) {
+                e.printStackTrace();
+            }
+            UnicastRemoteObject.unexportObject(this, true);
+            System.out.println("Game over.");
         }
-        UnicastRemoteObject.unexportObject(this, true);
-        System.out.println("Game over.");
+        return gameOver;
     }
 
     /** Paste the current game on screen. */

### 2017-11-22 GameInt.java

@@ -21,5 +21,4 @@ public interface GameInt extends Remote {
 
     public void printBoard() throws RemoteException;
 
-    public void endGame() throws RemoteException;
 }

### 2017-11-22 Leader.java

@@ -62,32 +62,19 @@ public class Leader extends Role {
         int play = maxIndices.get(new Random().nextInt(maxIndices.size()));
 
         leader.broadcastPlay(play);
-        broadcastPlay(play);
 
-        // Necessary because of the current implementation for ending the game.
-        // Might be unnecessary in the future.
-        if (leader != null)
+        // Broadcast the play. If the game is not over, start a new turn
+        if (!broadcastPlay(play))
             leader.turnStarts();
     }
 
     @Override
-    public synchronized void broadcastPlay(int play) throws RemoteException {
+    public synchronized boolean broadcastPlay(int play) throws RemoteException {
         // TODO: treat winning and losing differently
-        // TODO: (optional) start a new game or something?
-        // FIXME: regular players don't exit
         boolean gameOver = false;
         for (GameInt player : myTeam)
             gameOver = player.makePlay(play);
-        if (gameOver) {
-            leader = null;
-            if (myTeam.size() > 1)
-                for (GameInt player : myTeam.subList(1, myTeam.size() - 1))
-                    player.endGame();
-
-            // End the local game last. TODO: I don't thing this matters. Test and simplify if possible.
-            myTeam.get(0).endGame();
-            myTeam = null;
-        }
+        return gameOver;
     }
 
 	public void schedule() {

### 2017-11-22 Player.java

@@ -50,4 +50,8 @@ public class Player extends Role {
             //raiseElection();
         }
     }
+
+    public void cancelTimer() {
+        timer.cancel();
+    }
 }

### 2017-11-22 Role.java

@@ -11,7 +11,8 @@ public abstract class Role {
     public abstract GameInt addPlayer(GameInt player) throws RemoteException;
 
     /** A leader overrides this, a regular player does nothing */
-    public void broadcastPlay(int play) throws RemoteException {
+    public boolean broadcastPlay(int play) throws RemoteException {
+        return false;
     }
 
     public void setLeader(GameInt somePlayer) {
@@ -22,5 +23,8 @@ public abstract class Role {
 
     /** Regular players ignore this. */
     public void turnStarts() throws RemoteException {
+    }
+
+    public void cancelTimer() {
     }
 }

### 2017-11-23 paper/paper.tex
Added paragraph in System Design about Bully algorithm, Added comment in Implementation for Bully Algorithm

--- 2017-11-20 paper/paper.tex
@@
 The leaders communicate the vote of their team to the opposing leader so that they may pass that result on to the peers of that team, as pictured in figure 1.
+
+By considering the fact that only the two leaders communicate about the votes between the two teams, then the system can easily break if one of the leaders creashes. 
+In order to avoid that and provide a more robust system we chose to implemet the Bully Algorithm on each of the players so they will be able to detect a leader crash if they do not receive a vote reply from them in a certain period of time.
+After that, they can easily call for an election on older processes than them and wait for a leader to be announced on a different period of time.
+If no leader is announced on that period of time or there are no older processes, then that process can declare itself leader to the rest of the team and the opponent leader.
+
 \section{Implementation}
+
+% Bully Implementation
+% All of the players keep track of their leader, the opponent leader and a priority list of their team.
+% In this case, they can easily use the list to start election on the processes older than them.
 

### 2017-11-23 paper/paper.tex

--- 2017-11-22 paper/paper.tex
@@
 The leaders communicate the vote of their team to the opposing leader so that they may pass that result on to the peers of that team, as pictured in figure 1.
+
+By considering the fact that only the two leaders communicate about the votes between the two teams, then the system can easily break if one of the leaders creashes. 
+In order to avoid that and provide a more robust system we chose to implemet the Bully Algorithm on each of the players so they will be able to detect a leader crash if they do not receive a vote reply from them in a certain period of time.
+After that, they can easily call for an election on older processes than them and wait for a leader to be announced on a different period of time.
+If no leader is announced on that period of time or there are no older processes, then that process can declare itself leader to the rest of the team and the opponent leader.
+
 \section{Implementation}
+
 The implementation follows the RMI distributed object model. A GameInt interface extends java.rmi.Remote and lists remote classes. 
@@
  which can then inform the members of their team.
+
+% Bully Implementation
+% All of the players keep track of their leader, the opponent leader and a priority list of their team.
+% In this case, they can easily use the list to start election on the processes older than them.
+
 \section{Evaluation}

### 2017-11-23 paper/paper.tex
Added NAT-traversal comments to paper as Architecture Evaluation section

@@
 By refraining from introducing additional complexity as part of the goal of the project, and 
-allowing the users to see the distributed nature of the software being demonstrated via a simple game, the focus is shifted towards the implications of the distributive aspects of the system and 
-the way they are handled.
+allowing the users to see the distributed nature of the software being demonstrated via a simple game, the focus is shifted towards the implications of the distributive aspects of the system and the way they are handled.
 The finished product can serve to entertain the end users, as well as increase their curiosity and insight about computers and distributed systems.
@@
 
-By considering the fact that only the two leaders communicate about the votes between the two teams, then the system can easily break if one of the leaders creashes. 
+By considering the fact that only the two leaders communicate about the votes between the two teams, then the system can easily break if one of the leaders crashes.
+
 In order to avoid that and provide a more robust system we chose to implemet the Bully Algorithm on each of the players so they will be able to detect a leader crash if they do not receive a vote reply from them in a certain period of time.
+
 After that, they can easily call for an election on older processes than them and wait for a leader to be announced on a different period of time.
+
 If no leader is announced on that period of time or there are no older processes, then that process can declare itself leader to the rest of the team and the opponent leader.
@@
 The implementation follows the RMI distributed object model. A GameInt interface extends java.rmi.Remote and lists remote classes.
-A Game class implements GameInt and extends the UnicastRemoteObject. It is an analogy for what a player knows and does during a game. 
-It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, 
-setting the leader of the game, several methods to interact with the Role of the player and a main method that connects to other game instances. 
-Currently, when running the game, the user passes it a local instance name and the name of another instance as arguments. Each instance then uses this information 
-to do an RMI Registry lookup and get a reference to the other client. The Role interface allows the application to differentiate between Leader and Player classes which extend it. 
+
+A Game class implements GameInt and extends the UnicastRemoteObject.
+It is an analogy for what a player knows and does during a game.
+It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, setting the leader of the game, several methods to interact with the Role of the player and a main method that connects to other game instances.
+C
+urrently, when running the game, the user passes it a local instance name and the name of another instance as arguments.
+Each instance then uses this information to do an RMI Registry lookup and get a reference to the other client.
+The Role interface allows the application to differentiate between Leader and Player classes which extend it.
+
 It utilises the State design pattern to allow players to seamlessly have their role changed – for example when the current leader disconnects.
-A Leader is responsible of collecting the votes from all of the members in their team in the turnStarts() method. This is done asynchronously using
-a separate thread for each member – an instance of VoteThread. Each thread collects a vote from one assigned player then writes it into a corresponding position
-in a shared AtomicIntegerArray that they have all been passed by the Leader. This ensures the main Leader thread can then read these values. 
+
+A Leader is responsible of collecting the votes from all of the members in their team in the turnStarts() method.
+This is done asynchronously using a separate thread for each member – an instance of VoteThread.
+Each thread collects a vote from one assigned player then writes it into a corresponding position in a shared AtomicIntegerArray that they have all been passed by the Leader.
+This ensures the main Leader thread can then read these values.
+
 Each Leader then informs all players in the game about the vote of his team. A Player instance represents a regular player that only must vote.
-Each regular player also has a mechanism to detect if a leader has not contacted them to choose a play for too long. If a time-out is triggered, 
-a player calls for a new leader election. To avoid many simultaneous election calls, each regular player waits 5 minutes plus a randomised amount of time.
-All plays are recorded on a GameState object instance that acts as a game board. After each turn it checks if the game is over and relays this message to the leaders
- which can then inform the members of their team.
+
+Each regular player also has a mechanism to detect if a leader has not contacted them to choose a play for too long.
+If a time-out is triggered, a player calls for a new leader election.
+To avoid many simultaneous election calls, each regular player waits 5 minutes plus a randomised amount of time.
+
+All plays are recorded on a GameState object instance that acts as a game board.
+After each turn it checks if the game is over and relays this message to the leaders which can then inform the members of their team.
 
@@
 
-\section{Evaluation}
+\section{Performance Evaluation}
+
+\section{Architecture Evaluation}
+While Java RMI provides a simplified approach to object manipulation over distributed systems, it does force the use of TCP sockets by default.
+The reliable, ordered delivery of data provided by TCP is desirable for distributed applications, however, this choice of transport protocol does have significant implications for application topology.
+
+The ongoing shortage of IPv4 addresses has led to a large proportion of machines on the Internet to be obscured by Network Address Translation (NAT) middleboxes.
+Machines hidden behind one of these middleboxes cannot be contacted directly without the use of techniques like hole punching, which is a significant problem for establishing an entirely decentralised peer-to-peer system.
+
+Hole punching for NAT traversal requires the use of an external, publicly-accessible proxy server to distribute the public-facing IP address and port combinations (addr:port) of NAT-obscured clients which want to establish a peer-to-peer session.
+This traversal method is used by many popular distributed applications such as multiplayer gaming, Skype, and peer-to-peer filesharing.
+This technique takes advantage of the behaviour of NAT middleboxes to create and maintain an 'open' addr:port combination - without this, a fully decentralised peer-to-peer session cannot be established.
+
+As mentioned above, Java RMI uses TCP for its operations by default.
+While it is possible to modify the sockets created by RMI calls to use UDP instead, this requires extensive work using RMISocketFactory which is beyond the time allowed by the project.
+
+UDP is the more commonly used transport protocol for peer-to-peer applications - one socket can be used for both sending and receiving messages from multiple clients without the need for establishing a fixed connection between two endpoints.
+Although adjustments would need to be made to allow for unordered or lost messages, UDP would allow us to implement a topology with good fault tolerance and Internet-wide access, as displayed in figure (INCLUDE IMAGE HERE).
+
+An alternative approach is to attempt TCP hole punching instead.
+This requires creating four TCP sockets with bound to the same local addr:port combination (ie.
+with SO\_REUSEADDR enabled) to establish a connection with a peer through a proxy server, as described in section 4 of a study by Ford, et al (INCLUDE CITATION HERE).
+Not only is this a significant overhead, with four sockets and multiple ports required for each peer a client wants to make contact with, it is also more complex to perform than the UDP equivalent; our simplified Python-based tests could not establish contact between peers using the instructions described in the Ford paper.
+
+Without significant adjustments to the type of sockets used by Java RMI, a better understanding of TCP hole punching, or the universal adoption of IPv6, our application is currently limited to operating within local networks only.
+This is acceptable for its use as an educational tool, where it can be used within corporate or university networks freely, but the lack of NAT traversal largely excludes its use as a casual multiplayer game given that the vast majority of homes obscure several devices behind an ISP-provided NAT middlebox.
+
+While we are aiming to create a system without unnecessary complexity for educational purposes, NAT is a core component of the modern Internet - all networks programmers will encounter difficulties with it at some stage and having a solution to demonstrate as part of our application would have been very beneficial to people using it for educational purposes.
 

### 2017-11-23 Game.java
Fixed bully algorithm

@@ -71,8 +71,8 @@ public class Game extends UnicastRemoteObject implements GameInt {
     }
 
     @Override
-    public boolean startElection() throws RemoteException {
-        return role.startElection();
+    public boolean startElection(GameInt thisGameInt) throws RemoteException {
+        return role.startElection(thisGameInt);
     }
 
     @Override
@@ -90,7 +90,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
 
     /** Make this instance into a leader and initialise the team list */
     public void setLeader(PriorityBlockingQueue<GameInt> team) throws RemoteException {
-        this.role = new Leader(team);
+        this.role = new Leader(team);  // This one should create a leader with its team, leader and opponent leader
         System.out.println("I am a leader now.");
     }
 

### 2017-11-23 GameInt.java

@@ -36,14 +36,15 @@ public interface GameInt extends Remote {
     // If player not oldest:
     //       if timeoutElection then consider itself coordinator
     //       if timeoutCoordinator then start another election
-    public int timeoutElection = 10, timeoutCoordinator = 10;
+    public int timeoutCoordinator = 10;
 
 
     // Use this to start an election from oldest processes than you
     // Response is used for the timeout in the calling process
-    public boolean startElection() throws RemoteException;
+    public boolean startElection(GameInt thisGameInt) throws RemoteException;
 
 
     // Use this to broadcast to all the processes in your team & opponent leader , about the new leader
     public void declareLeader(GameInt leader) throws RemoteException;
+
 }

### 2017-11-23 Leader.java

@@ -12,7 +12,7 @@ public class Leader extends Role {
     // private List<GameInt> myTeam; --> moved to Role
 
     public Leader(PriorityBlockingQueue<GameInt> team) {
-        this.team = team;
+        super(team);
     }
 
     public Leader(GameInt leader,GameInt opponentLeader,PriorityBlockingQueue<GameInt> team) {

### 2017-11-23 Role.java

@@ -53,25 +53,32 @@ public abstract class Role{
     // False indicates leader has no "greater" leaders
     // True indicates leader has started another election
     // No response assume other process crashed -- Use timeout when calling it
-    public boolean startElection() throws RemoteException {
+    public boolean startElection(GameInt myGameInt) throws RemoteException {
         // Start an election to my team
         GameInt[] myTeam = (GameInt[]) this.team.toArray();
         ArrayList<Boolean> responses = new ArrayList<>();
+        responses.add(false);                                // No other leader
         int currentGameInt = 0;
 
-        while(this.hashCode() < myTeam[currentGameInt].hashCode() && (currentGameInt<myTeam.length)){
-            responses.add(myTeam[currentGameInt].startElection());
+        while(myGameInt.hashCode() < myTeam[currentGameInt].hashCode() && (currentGameInt<myTeam.length)){
+            boolean anotherLeaderExists = responses.remove(0);
+            responses.add(myTeam[currentGameInt].startElection(myTeam[currentGameInt]) && anotherLeaderExists);
             currentGameInt++;
         }
-        // No more leaders set me as leader
-        if (responses.size() == 0){
-            return false;
+        // No more leaders set me as leader to the rest and the leader
+        if (responses.get(0) == false){
+            for (GameInt player: myTeam) {
+                player.declareLeader(myGameInt);
+            }
+            opponentLeader.declareLeader(myGameInt);
         }
-        return true;
+
+        return responses.get(0);
     }
 
 
     public void declareLeader(GameInt leader) throws RemoteException {
         this.leader = leader;
     }
+
 }

### 2017-11-23 Role.java
Removed unecessary parts

@@ -32,15 +32,6 @@ public abstract class Role{
     /** Returns true if the added player is a leader */
     public abstract boolean addPlayer(GameInt player) throws RemoteException;
 
-    /** Returns true if the added player is an opponent leader*/
-    public synchronized boolean addOpponent(GameInt player) throws RemoteException{
-        if (opponentLeader == null) {
-            opponentLeader = player;
-            return true;
-        }
-        team.add(player);
-        return false;
-    }
 
     /** A leader overrides this, a regular player does nothing */
     public void broadcastPlay(int play) throws RemoteException {

### 2017-11-23 GameInt.java
Unnecessary changes

@@ -15,8 +15,6 @@ public interface GameInt extends Remote {
 
     public void setLeader() throws RemoteException;
 
-//    public void setLeader(GameInt leader,GameInt opponent) throws RemoteException;
-
     public void turnStarts() throws RemoteException;
 
     public void printBoard() throws RemoteException;

### 2017-11-23 Role.java

@@ -32,7 +32,6 @@ public abstract class Role{
     /** Returns true if the added player is a leader */
     public abstract boolean addPlayer(GameInt player) throws RemoteException;
 
-
     /** A leader overrides this, a regular player does nothing */
     public void broadcastPlay(int play) throws RemoteException {
     }

### 2017-11-25 ElectionRaiser.java
Making some headway towards DC-friendliness

@@ -1,0 +1,20 @@
+import java.rmi.RemoteException;
+import java.util.TimerTask;
+
+/** Task that is scheduled by the timer */
+public class ElectionRaiser extends TimerTask {
+
+    private GameInt client;
+
+    public ElectionRaiser(GameInt client) {
+        this.client = client;
+    }
+
+    public void run() {
+        try {
+            client.startElection();
+        } catch (RemoteException e) {
+            e.printStackTrace();
+        }
+    }
+}

### 2017-11-25 Game.java

@@ -120,6 +120,10 @@ public class Game extends UnicastRemoteObject implements GameInt {
         team.add(player);
     }
 
+    public void removeFromTeam(GameInt player) throws RemoteException {
+        team.remove(player);
+    }
+
     /** Role access classes */
     public boolean addPlayer(GameInt player) throws RemoteException {
         return role.addPlayer(player);
@@ -132,26 +136,29 @@ public class Game extends UnicastRemoteObject implements GameInt {
     // False indicates leader has no "greater" leaders
     // True indicates leader has started another election
     // No response assume other process crashed -- Use timeout when calling it
-    public boolean startElection(GameInt myGameInt) throws RemoteException {
+    public boolean startElection() throws RemoteException {
         // Start an election to my team
-        GameInt[] myTeam = (GameInt[]) team.toArray();
+        System.out.println("Election in progress.");
+        team.remove(leader);
+        GameInt[] myTeam = team.toArray(new GameInt[team.size()]);
         ArrayList<Boolean> responses = new ArrayList<>();
         responses.add(false); // No other leader
         int currentGameInt = 0;
 
-        while (myGameInt.hashCode() < myTeam[currentGameInt].hashCode() && currentGameInt < myTeam.length) {
+        while (hashCode() < myTeam[currentGameInt].hashCode() && currentGameInt < myTeam.length) {
             boolean anotherLeaderExists = responses.remove(0);
-            responses.add(myTeam[currentGameInt].startElection(myTeam[currentGameInt]) && anotherLeaderExists);
+            responses.add(myTeam[currentGameInt].startElection() && anotherLeaderExists);
             currentGameInt++;
         }
 
         // No more leaders set me as leader to the rest and the leader
         if (responses.get(0) == false) {
             for (GameInt player : myTeam)
-                player.setLeader(myGameInt);
-            opponentLeader.setLeader(myGameInt);
+                player.setLeader(this);
+            opponentLeader.setOpponentLeader(this);
         }
 
+        System.out.println("Election finished.");
         return responses.get(0);
     }
 

### 2017-11-25 GameInt.java

@@ -32,11 +32,13 @@ public interface GameInt extends Remote {
 
     public void addToTeam(GameInt player) throws RemoteException;
 
+    public void removeFromTeam(GameInt player) throws RemoteException;
+
     public void setAsPlayer() throws RemoteException;
 
     public void turnStarts() throws RemoteException;
 
     public void printBoard() throws RemoteException;
 
-    public boolean startElection(GameInt thisGameInt) throws RemoteException;
+    public boolean startElection() throws RemoteException;
 }

### 2017-11-25 Leader.java

@@ -32,7 +32,7 @@ public class Leader extends Role {
         int teamSize = team.size();
         voteCollectors = new VoteThread[teamSize];
         for (GameInt player : team)
-            voteCollectors[counter++] = new VoteThread(player, votes);
+            voteCollectors[counter++] = new VoteThread(this, player, votes);
 
         for (VoteThread thread : voteCollectors)
             thread.start();
@@ -61,8 +61,13 @@ public class Leader extends Role {
         // Broadcast the play. If the game is not over, start a new turn
         GameInt opponentLeader = game.getOpponentLeader();
         opponentLeader.broadcastPlay(play);
-        if (!broadcastPlay(play))
-            opponentLeader.turnStarts();
+        if (!broadcastPlay(play)) {
+            try {
+                opponentLeader.turnStarts();
+            } catch (RemoteException e) {
+                // let the teammates handle this
+            }
+        }
     }
 
     @Override
@@ -73,8 +78,14 @@ public class Leader extends Role {
         return gameOver;
     }
 
-	public void schedule() {
-		// Schedule Timer to record last response of each regular player on team
-		// If idle for more than x minutes, delete from team and inform other members to remove reference
-	}
+    public void removeFromTeam(GameInt player) throws RemoteException {
+        game.removeFromTeam(player);
+        for (GameInt teamMember : game.getTeam())
+            teamMember.removeFromTeam(player);
+    }
+
+    public void schedule() {
+        // Schedule Timer to record last response of each regular player on team
+        // If idle for more than x minutes, delete from team and inform other members to remove reference
+    }
 }

### 2017-11-25 Player.java

@@ -13,6 +13,7 @@ public class Player extends Role {
     public Player(GameInt game) {
         super(game);
         timer = new Timer();
+        schedule();
     }
 
     public boolean addPlayer(GameInt player) throws RemoteException {
@@ -27,7 +28,8 @@ public class Player extends Role {
         Calendar cal=Calendar.getInstance(); // get today
         int hr = cal.get(Calendar.HOUR_OF_DAY);
         int mn = cal.get(Calendar.MINUTE);
-        mn = 5 * (mn / 5 + 1);	// round up to next 5 minute slot
+        //mn = 5 * (mn / 5 + 1);	// round up to next 5 minute slot
+        mn += 1;
         if (mn == 60) {
             hr = hr + 1;
             mn = 0;
@@ -40,15 +42,8 @@ public class Player extends Role {
         Random rand = new Random();
         int randInt = rand.nextInt(500) + 1;
         period = 5 * 60 * 1000 + randInt; // 5 minutes in milliseconds plus a random extension from 1 to 500ms
-        task = new ElectionRaiser();
+        task = new ElectionRaiser(game);
         timer.schedule(task, delay, period);
-    }
-
-    /** Task that is scheduled by the timer */
-    private class ElectionRaiser extends TimerTask {
-        public void run() {
-            //raiseElection();
-        }
     }
 
     public void cancelTimer() {

### 2017-11-25 VoteThread.java

@@ -3,10 +3,12 @@ import java.util.concurrent.atomic.AtomicIntegerArray;
 
 /** Gets a vote from a player. */
 class VoteThread extends Thread {
+    private Leader leader;
     private GameInt player;
     private AtomicIntegerArray votes;
 
-    public VoteThread(GameInt player, AtomicIntegerArray votes) {
+    public VoteThread(Leader leader, GameInt player, AtomicIntegerArray votes) {
+        this.leader = leader;
         this.player = player;
         this.votes = votes;
     }
@@ -15,7 +17,11 @@ class VoteThread extends Thread {
         try {
             votes.getAndIncrement(player.askForInput());
         } catch (RemoteException e) {
-            e.printStackTrace();
+            try {
+                leader.removeFromTeam(player);
+            } catch (RemoteException f) {
+                f.printStackTrace(); // the first exception is handled, this one is not
+            }
         }
     }
 }

### 2017-11-29 Game.java
Salvaging the election functionality to have something presentable for tomorrow

@@ -151,6 +151,7 @@ public class Game extends UnicastRemoteObject implements GameInt {
             currentGameInt++;
         }
 
+        this.setLeader();
         // No more leaders set me as leader to the rest and the leader
         if (responses.get(0) == false) {
             for (GameInt player : myTeam)

### 2017-11-29 Player.java

@@ -36,7 +36,7 @@ public class Player extends Role {
         }
         cal.set(Calendar.HOUR_OF_DAY, hr);
         cal.set(Calendar.MINUTE, mn);
-        cal.set(Calendar.SECOND, 0);
+        cal.set(Calendar.SECOND, cal.get(Calendar.SECOND));
         // milliseconds until next 5 minute boundary
         delay = cal.getTimeInMillis() - System.currentTimeMillis();
         Random rand = new Random();

### 2017-11-30 paper/paper.tex
architecture evaluation and related work revised

@@
 Hole punching for NAT traversal requires the use of an external, publicly-accessible proxy server to distribute the public-facing IP address and port combinations (addr:port) of NAT-obscured clients which want to establish a peer-to-peer session.
-This traversal method is used by many popular distributed applications such as multiplayer gaming, Skype, and peer-to-peer filesharing.
+This traversal method is used by many popular types of distributed applications such as multiplayer gaming, VoIP, and peer-to-peer file sharing.
 This technique takes advantage of the behaviour of NAT middleboxes to create and maintain an 'open' addr:port combination - without this, a fully decentralised peer-to-peer session cannot be established.
@@
 UDP is the more commonly used transport protocol for peer-to-peer applications - one socket can be used for both sending and receiving messages from multiple clients without the need for establishing a fixed connection between two endpoints.
-Although adjustments would need to be made to allow for unordered or lost messages, UDP would allow us to implement a topology with good fault tolerance and Internet-wide access, as displayed in figure (INCLUDE IMAGE HERE).
+Although adjustments would need to be made to allow for unordered or lost messages, UDP would allow us to implement a topology with robust fault tolerance and Internet-wide access:
 
-An alternative approach is to attempt TCP hole punching instead.
-This requires creating four TCP sockets with bound to the same local addr:port combination (ie.
-with SO\_REUSEADDR enabled) to establish a connection with a peer through a proxy server, as described in section 4 of a study by Ford, et al (INCLUDE CITATION HERE).
-Not only is this a significant overhead, with four sockets and multiple ports required for each peer a client wants to make contact with, it is also more complex to perform than the UDP equivalent; our simplified Python-based tests could not establish contact between peers using the instructions described in the Ford paper.
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{images/p2p-topology-3point.png}
+	\caption{Proposed Topology for UDP-based Version of System}
+	\label{fig:topology2}
+\end{figure}
+
+An alternative approach is to attempt TCP hole punching.
+This requires creating four TCP sockets bound to the same local addr:port combination (ie.
+with SO\_REUSEADDR enabled) to establish a connection with a peer through a proxy server, as described in section 4 of a paper by Ford, et al \cite{ford_srisuresh_kegel_2005}.
+Not only is this a significant overhead, with four sockets required for each peer a client wants to make contact with, it is also more complex to perform than the UDP equivalent - our initial tests could not establish contact between peers using the instructions described in the Ford paper.
 
 Without significant adjustments to the type of sockets used by Java RMI, a better understanding of TCP hole punching, or the universal adoption of IPv6, our application is currently limited to operating within local networks only.
-This is acceptable for its use as an educational tool, where it can be used within corporate or university networks freely, but the lack of NAT traversal largely excludes its use as a casual multiplayer game given that the vast majority of homes obscure several devices behind an ISP-provided NAT middlebox.
+This is acceptable for its use as an educational tool, where it can be used within corporate or university networks freely, but the lack of NAT traversal excludes its use as a casual multiplayer game given that the vast majority of homes obscure several devices behind an ISP-provided NAT middlebox.
 
-While we are aiming to create a system without unnecessary complexity for educational purposes, NAT is a core component of the modern Internet - all networks programmers will encounter difficulties with it at some stage and having a solution to demonstrate as part of our application would have been very beneficial to people using it for educational purposes.
+While we are aiming to create a system without unnecessary complexity, NAT is a core component of the modern Internet - all networks programmers will encounter difficulties with it at some stage, and having a solution to demonstrate as part of our application would have been very beneficial extra material for people using it for educational purposes.
 
 \section{Related Work}
-While other multiplayer distributed games already exist, they have not been
-created with this project’s task in mind and therefore do not serve the same
-purpose. They can be split into two categories – very unprofessional and too
-professional. For example, hobby implementations exist that deal with
-distributed problems in a bad fashion, promoting bad and incorrect programming 
-practises. Similarly, a single-player-per-team distributed tic-tac-toe game in 
-Java can be found online that does not comply with any design guidelines and 
-puts all functionality in a single class and as such, cannot teach students good 
-practice and has the potential of confusing the learner further about distributive 
-programming \cite{MrWayFarOutYoutube15}.
+While other distributed multiplayer games already exist, they do not serve the same
+purpose as our project.
+Existing products can be split into two categories: too simple and too complex.
+Examples of the former generally deal with the difficulties of distributed systems in a bad fashion, promoting questionable design and programming practices.
+One specific example of this is a YouTube guide to building a single player-per-team distributed tic-tac-toe game in Java \cite{MrWayFarOutYoutube15} - while the material is accessible for inexperienced networks programmers, it puts all functionality in a single class and does not comply with any design guidelines.
+As such, this resource has the potential to further confuse learners about programming distributed systems when they try to read more complex documentation which does obey software engineering design principles.
 
-Separately, very in-depth materials that provide state-of-the-art practices can 
-be found \cite{DBLP:conf/nsdi/BharambePS06} but there is a high possibility that they might be incomprehensible 
-to a beginner reader looking to tackle basic distributed concepts to begin with. 
-Furthermore, none of the similar implementations residing on the Internet take advantage 
-of the Java RMI API which must be considered a good starting point for distributed systems
-learners for it to be taught as part of our degree and is a requirement for this
-project. Instead, they use different sets of other technologies which can
-further confuse their reader and introduce a steeper learning curve.
+The latter category consists of resources which provide details of state-of-the-art practices and advanced optimisations.
+An example of this is a version of the game Quake II, modified to run multiplayer sessions more efficiently over a distributed system \cite{DBLP:conf/nsdi/BharambePS06}.
+This is excellent material for an experienced programmer looking to refine their skills, but daunting to a newcomer due to a high level of complexity in both networking concepts and game logic.
+
+Furthermore, none of the similar implementations residing on the Internet make use of the Java RMI API.
+Many distributed systems projects use other technologies, such as C or Rust, which have a far steeper learning curve than Java, and are not as commonly taught in universities or used as frequently in enterprise environments.
+
+In addition to being based on a widely-known language, Java RMI has the additional benefit of simplifying processes such as marshalling and socket creation.
+Our project takes advantage of this to allow people to learn about concepts in distributed systems programming without worrying about added complexities such as objects becoming corrupted due to incorrect formatting, or incorrect use of transport protocols.
 
 \section{Conclusions and Future Work}
+%make the bully algorithm work when both leaders crash at the same time
+%team balancing
+%randomised moves not implemented
 

### 2017-12-01 paper/paper.tex
Paper - Conclusion & Future Work

@@
 While previous implementations attempting to serve a similar purpose exist, none of them boast good design or efficiency and neither of them deliver the same functionality.
-The developed piece of software allows clients to play locally or globally a game of the classic tic tac toe - but with a twist.
+The developed piece of software allows clients to play a game of Tic Tac Toe locally - but with a twist.
 Players are separated into teams and decide on what the next move should be by voting.
@@
 
-The leaders communicate the vote of their team to the opposing leader so that they may pass that result on to the peers of that team, as pictured in figure 1.
+Upon initialising the game, the player will take a note of their address and port number.
+As the game is played on a local network, the player will attempt to connect to any other existing nodes on the network.
+If none are found, then the client will initialise a new game instance and set itself as one of the leaders.
 
-By considering the fact that only the two leaders communicate about the votes between the two teams, then the system can easily break if one of the leaders crashes.
+When other players attempt to check for existing players on the network and successfully find them, it will get the current game state, check if it is one of the leaders, print the board in text format and prepare for the turn.
 
-In order to avoid that and provide a more robust system we chose to implemet the Bully Algorithm on each of the players so they will be able to detect a leader crash if they do not receive a vote reply from them in a certain period of time.
-
-After that, they can easily call for an election on older processes than them and wait for a leader to be announced on a different period of time.
-
-If no leader is announced on that period of time or there are no older processes, then that process can declare itself leader to the rest of the team and the opponent leader.
+The leaders update a common game state at the end of each event before broadcasting the game state to all team members. Players are told by the leaders when completion of the game occurs.
 
@@
 It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, setting the leader of the game, several methods to interact with the Role of the player and a main method that connects to other game instances.
-C
-urrently, when running the game, the user passes it a local instance name and the name of another instance as arguments.
+
+Currently, when running the game, the user passes it a local instance name and the name of another instance as arguments.
 Each instance then uses this information to do an RMI Registry lookup and get a reference to the other client.
@@
 %randomised moves not implemented
+The produced software serves as a good demonstration of a distributed game.
+The use of a simple, well known game idea turned into a distributed, multiplayer experience highlights the benefits and complexities associated with distributed systems, and allows the technology to be demonstrated in an inclusive format, allowing users of all levels to make use of the software without additional complications added.
+
+For future additions or design reconsiderations, Distributed Tic Tac Toe could benefit from some changes.
+These changes include making the software more robust and granting the users an improved playing experience by addressing certain network issues and gameplay elements.
+
+In the softwares current state, leaders are appointed using the bully algorithm when a new election is required.
+However, there is no support for failure of communication from both leaders, i.e.
+if both leaders were to crash at the same time.
+Although the likelihood is slim, addressing the possibility of this scenario playing out would make the system more robust and prevent frustration for the players.
+
+To improve the gameplay experience, players of both teams should be on an even playing field.
+To ensure this is consistent throughout the duration of a match, some form of automatic team balancing is recommended, allowing each team to receive votes in as even amount as possible.
+The gameplay experience could be improved further still by addressing the scenario of an even number of votes being cast by team members.
+The possibility of this happening is certainly high, and the recommendation is that the leader randomly selects between the moves chosen by the team members.
 

### 2017-12-01 paper/paper.tex
Futher revisions to paper for readability

@@
 \begin{abstract}
-This article describes a Distributed Multiplayer Tic Tac Toe software game written in Java using the RMI API.
-While previous implementations attempting to serve a similar purpose exist, none of them boast good design or efficiency and neither of them deliver the same functionality.
-The developed piece of software allows clients to play a game of Tic Tac Toe locally - but with a twist.
-Players are separated into teams and decide on what the next move should be by voting.
-The implementation is derived from a design that follows the distributed applications guidelines and is prepared to handle corner cases such as client disconnects for example.
+This article describes a Distributed Multiplayer Tic-Tac-Toe software game written in Java using the RMI API.
+While previous implementations attempting to serve a similar purpose exist, none of them boast good design or efficiency, and neither of them deliver the same functionality.
+This project allows clients to play a game of Tic-Tac-Toe locally, with a twist: players collecively decide on what the next move should be by voting in teams.
+The implementation is derived from a design that follows the distributed applications guidelines and is prepared to handle corner cases such as client disconnects.
 \end{abstract}
@@
 \section{Introduction}
-Coding is the main occupation of many people nowadays, and it is a
-past-time for many too. Most beginners get their first steps in writing code
-following the local single-threaded model. Often, it can be difficult for
-new programmers, or even those with more experience, to abstract the key
-knowledge they have obtained using this model and then apply it in a different way.
+Programming is very much an in-demand profession in the current job market, as well as being a popular hobby. Most beginners take their first steps in writing code
+following a simple, single-threaded model with zero distribution.
+It is often difficult for new programmers to abstract the key knowledge they have obtained to a point where they can independently apply it to a different scenario.
 
-Distributed software requires a new look on the system – the number of devices
-has gone up and with it so have the number of processes, concerns and complexities.
-Some concepts have now changed and need to be dealt with differently; One such
-example if this would be mutual exclusion. We believe programs that serve as examples and
-focus on simple local tasks turned into distributed ones can help many people
-trying to make this step with a hands-on experience of what they are preparing
-for. Seeing theory in practice is a proven technique and there is evidence for
-this in every university course that features practical work. This is especially
-true where games are concerned, proving to be an effective means of communicating 
-programming methods and techniques in an engaging way
-\cite{Kiniry:2011:VG:1984674.1984681}. For this reason, the team has decided to
-take a simple, well known game and develop it into a distributed system; Tic Tac
-Toe played as a peer-to-peer, team-based gaming system where the players are
-organised into teams and required to vote on moves for their team.
+Distributed software engineering principles are evolving quickly – the number of Internet-enabled devices has rapidly increased in recent years, and the level of complexity in creating programs to use these interconnected systems effectively has increased along with it.
+This complexity arises from core concepts such as co-ordinating resource sharing and fault tolerance - these are challenging concepts to learn for a novice programmer, but they are integral to creating a reliable and effective distributed system.
 
-By refraining from introducing additional complexity as part of the goal of the project, and 
-allowing the users to see the distributed nature of the software being demonstrated via a simple game, the focus is shifted towards the implications of the distributive aspects of the system and the way they are handled.
-The finished product can serve to entertain the end users, as well as increase their curiosity and insight about computers and distributed systems.
-Using Tic Tac Toe as a foundation to demonstrate a distributed system benefits from both the simplicity and familiarity of the game itself, meaning no complex rulesets or in-depth mechanisms need be explained to the users, allowing the distributed nature to be a novel way of experiencing a widely known game.
+We believe that programs which serve as examples of turning familiar, relatively simple local applications into distributed ones are key to bridging this gap, providing a hands-on introduction to a field which is becoming ever more important with the deployment of novel concepts like fog computing.
+The sheer number of programming tutorials available on the internet shows that there is a demand for model solutions for educational purposes.
+Games in particular have proven to be an effective means of communicating programming methods and techniques in an engaging way \cite{Kiniry:2011:VG:1984674.1984681}.
+For this reason, we have decided to take a simple, well known game and develop it into a distributed system: Tic-Tac-Toe played as a peer-to-peer, team-based gaming system where the players are organised into teams and required to vote the moves their team will make.
+
+By refraining from introducing unnecessary complexity into the project, and allowing the users to see an example of distributed software being demonstrated via a simple game, the focus is shifted towards the implications of the distributive aspects of the system and the way they are handled.
+Using Tic-Tac-Toe as a foundation to demonstrate a distributed system benefits from both the simplicity and familiarity of the game itself, meaning no complex rulesets or in-depth mechanisms need be explained to the users, allowing the distributed nature to be a novel way of experiencing a widely known game.
 
@@
 %team balancing
-%randomised moves not implemented
+
 The produced software serves as a good demonstration of a distributed game.
-The use of a simple, well known game idea turned into a distributed, multiplayer experience highlights the benefits and complexities associated with distributed systems, and allows the technology to be demonstrated in an inclusive format, allowing users of all levels to make use of the software without additional complications added.
+The use of a simple, well known game turned into a distributed, multiplayer experience highlights the benefits and complexities associated with distributed systems, and allows the technology to be demonstrated in an accessible format, allowing users of all levels to make use of the software without being discouraged by an overwhelming level of complexity.
 
-For future additions or design reconsiderations, Distributed Tic Tac Toe could benefit from some changes.
-These changes include making the software more robust and granting the users an improved playing experience by addressing certain network issues and gameplay elements.
+Distributed Tic-Tac-Toe could benefit from some changes in future iterations, including making the software more robust and granting the users an improved playing experience by addressing certain network issues and gameplay elements.
 
-In the softwares current state, leaders are appointed using the bully algorithm when a new election is required.
+In the project's current state, leaders are appointed using the bully algorithm when one machine acting as a leader fails.
 However, there is no support for failure of communication from both leaders, i.e.
 if both leaders were to crash at the same time.
-Although the likelihood is slim, addressing the possibility of this scenario playing out would make the system more robust and prevent frustration for the players.
+Although the likelihood of this occurring is slim, this possibility still needs to be addressed to improve player experience and to improve the standard of our product as an educational resource.
 
-To improve the gameplay experience, players of both teams should be on an even playing field.
-To ensure this is consistent throughout the duration of a match, some form of automatic team balancing is recommended, allowing each team to receive votes in as even amount as possible.
-The gameplay experience could be improved further still by addressing the scenario of an even number of votes being cast by team members.
-The possibility of this happening is certainly high, and the recommendation is that the leader randomly selects between the moves chosen by the team members.
+Another improvement which we would make in further iterations would be some form of automatic team balancing - currently, no replacement team members are allocated to an underpopulated team if several peers drop out of a session.
+Another situation which needs to be addressed is the scenario of all members of a team leaving a game before it is complete, an event which currently has no resolution.
+
+The simplicity of Tic-Tac-Toe, combined with NAT traversal issues, limits our project's appeal as a multiplayer game to people with no interest in programming.
+However, our commitment to obeying software engineering design principles in Java makes this a valuable educational tool for beginners wanting hands-on experience with engineering a distributed, peer-to-peer application.
 

### 2017-12-01 paper/paper.tex
more diagrams and clarifications added

@@
 \section{System Design}
-Tic Tac Toe is a simple game played on a grid of 3x3, consisting of two players.
-One player uses an “X” marker and the other player uses an “O” marker to select a grid space to occupy.
+Tic-Tac-Toe is a simple game played on a 3x3 grid, consisting of two players.
+One player uses an “X” marker and the other player uses an “O” marker occupy spaces in the grid.
 The objective of the game is to take turns until one player manages to occupy three grid spaces in a straight line, be it horizontal, vertical or diagonal.
 
-The proposed design takes this simple concept and uses it to demonstrate a distributed system, whereby the two “players” are instead two teams of up to 5 players, which take turns to vote on where their team should place their markers.
-Each team has a designated "leader" which communicates data with their team's peers, and the opposing team's "leader" to synchronise current gameplay state.
+The proposed design takes this simple concept and uses it to demonstrate a distributed system, whereby each player is assigned to one of two teams, each of which takes turns to vote on where their team should place their markers.
+Each team has a designated leader which communicates data with their team's peers, and the opposing team's leader to synchronise current gameplay state.
 
@@
 Upon initialising the game, the player will take a note of their address and port number.
-As the game is played on a local network, the player will attempt to connect to any other existing nodes on the network.
+As the game is played on a local network, the player will attempt to connect to all other existing nodes on the network.
 If none are found, then the client will initialise a new game instance and set itself as one of the leaders.
 
-When other players attempt to check for existing players on the network and successfully find them, it will get the current game state, check if it is one of the leaders, print the board in text format and prepare for the turn.
+If a player manages to connect to another machine on the local network (ie.
+that machine is paricipating in an active game session), they will retrieve the current game state, check if they are one of the leaders, print the current board state, and prepare for the turn.
 
-The leaders update a common game state at the end of each event before broadcasting the game state to all team members. Players are told by the leaders when completion of the game occurs.
+The leaders update a common game state at the end of each event before broadcasting this updated game state to all team members.
+Players are told by the leaders when the game, and therefore the session, ends.
 
@@
 It is an analogy for what a player knows and does during a game.
-It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, setting the leader of the game, several methods to interact with the Role of the player and a main method that connects to other game instances.
+It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, setting the leader of the game, several methods to interact with the Role of the player, and a main method that connects to other game instances.
 
-Currently, when running the game, the user passes it a local instance name and the name of another instance as arguments.
+Currently, when running the game, the user supplies a local instance name and the name of another instance as command line arguments.
 Each instance then uses this information to do an RMI Registry lookup and get a reference to the other client.
+
 The Role interface allows the application to differentiate between Leader and Player classes which extend it.
+It utilises the State design pattern to allow players to seamlessly have their role changed when required (eg.
+if the current leader disconnects).
 
-It utilises the State design pattern to allow players to seamlessly have their role changed – for example when the current leader disconnects.
+A Leader is responsible for collecting the votes from all of the members in their team in the turnStarts() method.
+This is done asynchronously using a separate thread for each member – an instance of VoteThread.
+Each thread collects a vote from one assigned player then writes it into a corresponding position in a shared AtomicIntegerArray that they have all been passed by the leader.
+This ensures the main leader thread can then read these values.
 
-A Leader is responsible of collecting the votes from all of the members in their team in the turnStarts() method.
-This is done asynchronously using a separate thread for each member – an instance of VoteThread.
-Each thread collects a vote from one assigned player then writes it into a corresponding position in a shared AtomicIntegerArray that they have all been passed by the Leader.
-This ensures the main Leader thread can then read these values.
-
-Each Leader then informs all players in the game about the vote of his team. A Player instance represents a regular player that only must vote.
+Each Leader then informs all players in the game about the most popular vote chosen by their team.
+In the event of two or more moves receiving an equal number of votes, a leader will randomly select  A Player instance represents a regular player that can only vote, it cannot independently generate and broadcast a new game state.
 
@@
 If a time-out is triggered, a player calls for a new leader election.
-To avoid many simultaneous election calls, each regular player waits 5 minutes plus a randomised amount of time.
+To avoid several simultaneous election calls, each regular player waits five minutes plus a smaller randomised amount of time.
 
-All plays are recorded on a GameState object instance that acts as a game board.
-After each turn it checks if the game is over and relays this message to the leaders which can then inform the members of their team.
+Elections use the bully algorithm to determine which Player instance will become the new Leader instance for their team.
+The other team members and the leader of the opposing team are all informed of the new leader, creating an altered topology to compensate for the failure:
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{images/DAS-topology-election.png}
+	\caption{Proposed Topology for UDP-based Version of System}
+	\label{fig:topology-election}
+\end{figure}
+
+If the failed machine joins the game session again, it will not automatically regain Leader status, instead participating as a regular Player instance.
+
+All plays are recorded on a GameState object instance which acts as a game board.
+After each turn, GameState checks if the game is over and relays this message to the leaders which can then inform the members of their team.
 
@@
 	\caption{Proposed Topology for UDP-based Version of System}
-	\label{fig:topology2}
+	\label{fig:topology-udp}
 \end{figure}

### 2017-12-01 paper/paper.tex
citations fixed

@@
 Currently, when running the game, the user supplies a local instance name and the name of another instance as command line arguments.
-Each instance then uses this information to do an RMI Registry lookup and get a reference to the other client.
+Each instance then uses this information to do an RMI Registry lookup and get a reference to the other peer to attempt to establish a session.
+This allows multiple different sessions to run simultaneously on the same network without interference - ideal for educational use, where multiple groups are likely to be working on the same internal network.
 

### 2017-12-01 createpeers.py
Created automated client generation script

@@ -1,0 +1,30 @@
+import sys
+import subprocess
+import time
+
+if sys.argv[1] == "server":
+    try:
+        #run rmiregistry in the background
+        subprocess.call("rmiregistry &", shell=True)
+        #create base server instance
+        subprocess.call("java Game server client localhost", shell=True)
+        print("Server created")
+    except Exception as e:
+        print(e)
+
+elif sys.argv[1] == "clients":
+    try:
+        #number of peers to create in the session
+        peer_count = int(sys.argv[2])
+        for i in range(peer_count):
+            #create instances named client0, client1, etc. in a new window
+            print("Creating peer %d" % i)
+            subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost\"" % i, shell=True)
+            #sleep to allow time for server to resolve connections
+            time.sleep(5)
+        print("%d peers created" % peer_count)
+    except Exception as e:
+        print(e)
+
+else:
+    print("Usage:\npython server - creates base server instance on localhost\npython clients [number] - creates [number] of clients to connect to server instance")

### 2017-12-01 createpeers.py
updated peer creation script

@@ -1,11 +1,16 @@
 import sys
 import subprocess
 import time
+import random
+import threading
+
+def client_thread(i, move):
+    #Create client, execute move on repeat
+    subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost; sleep 5; while true; do %d; done\"" % (i, move), shell=True)
+    
 
 if sys.argv[1] == "server":
     try:
-        #run rmiregistry in the background
-        subprocess.call("rmiregistry &", shell=True)
         #create base server instance
         subprocess.call("java Game server client localhost", shell=True)
         print("Server created")
@@ -17,9 +22,12 @@ elif sys.argv[1] == "clients":
         #number of peers to create in the session
         peer_count = int(sys.argv[2])
         for i in range(peer_count):
-            #create instances named client0, client1, etc. in a new window
-            print("Creating peer %d" % i)
-            subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost\"" % i, shell=True)
+            #create instance threads named client0, client1, etc. in a new window
+            print("Creating peer thread %d" % i)
+            move = random.randint(0,8)
+            c_thread = threading.Thread(target=client_thread, args=(i, move))
+            c_thread.start()
+            #subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost\"" % i, shell=True)
             #sleep to allow time for server to resolve connections
             time.sleep(5)
         print("%d peers created" % peer_count)
@@ -27,4 +35,4 @@ elif sys.argv[1] == "clients":
         print(e)
 
 else:
-    print("Usage:\npython server - creates base server instance on localhost\npython clients [number] - creates [number] of clients to connect to server instance")
+    print("Usage:\npython server - creates base server instance on localhost\npython clients [number] creates [number] of clients to connect to server instance")

### 2017-12-01 ElectionRaiser.java
More comments

@@ -1,7 +1,7 @@
 import java.rmi.RemoteException;
 import java.util.TimerTask;
 
-/** Task that is scheduled by the timer */
+/** Task scheduled by the timer to start an election */
 public class ElectionRaiser extends TimerTask {
 
     private GameInt client;

### 2017-12-01 GameState.java

@@ -1,5 +1,6 @@
 import java.io.Serializable;
 
+/** Represents the state of the game. Each player has an instance of this class. */
 public class GameState implements Serializable {
 
     /** Serial version UID to avoid warning. */

### 2017-12-01 Leader.java

@@ -70,6 +70,7 @@ public class Leader extends Role {
         }
     }
 
+    /** Send information about a player to all members of my team (including myself) */
     @Override
     public boolean broadcastPlay(int play) throws RemoteException {
         boolean gameOver = false;
@@ -78,14 +79,10 @@ public class Leader extends Role {
         return gameOver;
     }
 
+    /** Remove a player from the team (in case of a detected disconnect) */
     public void removeFromTeam(GameInt player) throws RemoteException {
         game.removeFromTeam(player);
         for (GameInt teamMember : game.getTeam())
             teamMember.removeFromTeam(player);
-    }
-
-    public void schedule() {
-        // Schedule Timer to record last response of each regular player on team
-        // If idle for more than x minutes, delete from team and inform other members to remove reference
     }
 }

### 2017-12-01 Player.java

@@ -5,6 +5,7 @@ import java.util.Timer;
 import java.util.TimerTask;
 import java.util.concurrent.PriorityBlockingQueue;
 
+/** A regular player who is not a leader. This class is only meant to hold functionality different from the Leader class */
 public class Player extends Role {
 
     private Timer timer;
@@ -20,6 +21,7 @@ public class Player extends Role {
         return game.getLeader().addPlayer(player);
     }
 
+    /** (Re)set the timer that waits to start an election */
     public void schedule() {
         if (task != null)
             task.cancel();

### 2017-12-01 Role.java

@@ -14,12 +14,11 @@ public abstract class Role {
         this.game = game;
     }
 
-    public abstract void schedule();
-
-    /** Returns a reference to the leader of the team or null if the added player is the new leader */
+    /** Return true if the added player becomes a leader */
     public abstract boolean addPlayer(GameInt player) throws RemoteException;
 
-    // Default implementations for regular players - avoids testing whether an instance is a leader or not
+  // Default empty implementations so that we don't have to check whether an instance of Role is actually
+  // Player or Leader
 
     public boolean broadcastPlay(int play) throws RemoteException {
         return false;
@@ -29,6 +28,9 @@ public abstract class Role {
     }
 
     public void cancelTimer() {
+    }
+  
+    public void schedule() {
     }
 
 }

### 2017-12-01 VoteThread.java

@@ -3,9 +3,9 @@ import java.util.concurrent.atomic.AtomicIntegerArray;
 
 /** Gets a vote from a player. */
 class VoteThread extends Thread {
-    private Leader leader;
-    private GameInt player;
-    private AtomicIntegerArray votes;
+    private Leader leader; // The calling class. Also the class we contact if the player fails
+    private GameInt player; // The player to get a vote from
+    private AtomicIntegerArray votes; // A reference to data structure that saves votes from all players
 
     public VoteThread(Leader leader, GameInt player, AtomicIntegerArray votes) {
         this.leader = leader;

### 2017-12-01 paper/paper.tex
Fixed some glaring mistakes in the paper. Recompile before submitting

--- 2017-12-01 paper/paper.tex
@@
 \begin{abstract}
-This article describes a Distributed Multiplayer Tic-Tac-Toe software game written in Java using the RMI API.
-While previous implementations attempting to serve a similar purpose exist, none of them boast good design or efficiency, and neither of them deliver the same functionality.
-This project allows clients to play a game of Tic-Tac-Toe locally, with a twist: players collecively decide on what the next move should be by voting in teams.
-The implementation is derived from a design that follows the distributed applications guidelines and is prepared to handle corner cases such as client disconnects.
+This article describes a distributed multiplayer Tic-Tac-Toe game
+written in Java using the RMI API. While previous implementations
+attempting to serve a similar purpose exist, none of them boast good
+design or efficiency, and neither of them deliver the same
+functionality. Our implementation allows clients to play a game of
+Tic-Tac-Toe with a twist: players collectively decide on what the next
+move should be by voting in teams. The implementation is derived from
+a design that follows the distributed applications guidelines and is
+prepared to handle corner cases such as client disconnects.
 \end{abstract}
 
-\section{Introduction}
+\section{Introduction} % TODO: cite the first 2 paragraphs
 Programming is very much an in-demand profession in the current job market, as well as being a popular hobby. Most beginners take their first steps in writing code
-following a simple, single-threaded model with zero distribution.
-It is often difficult for new programmers to abstract the key knowledge they have obtained to a point where they can independently apply it to a different scenario.
+with no concurrency. It is often difficult for new programmers to abstract the key knowledge they have obtained to a point where they can independently apply it to a different scenario.
 
-Distributed software engineering principles are evolving quickly – the number of Internet-enabled devices has rapidly increased in recent years, and the level of complexity in creating programs to use these interconnected systems effectively has increased along with it.
-This complexity arises from core concepts such as co-ordinating resource sharing and fault tolerance - these are challenging concepts to learn for a novice programmer, but they are integral to creating a reliable and effective distributed system.
+Distributed software engineering principles are evolving quickly --- the number of Internet-enabled devices has rapidly increased in recent years, and the level of complexity in creating programs to use these interconnected systems effectively has increased along with it.
+This complexity arises from core concepts such as co-ordinating resource sharing and fault tolerance --- these are challenging concepts to learn for a novice programmer, but they are integral to creating a reliable and effective distributed system.
 
@@
 Games in particular have proven to be an effective means of communicating programming methods and techniques in an engaging way \cite{Kiniry:2011:VG:1984674.1984681}.
-For this reason, we have decided to take a simple, well known game and develop it into a distributed system: Tic-Tac-Toe played as a peer-to-peer, team-based gaming system where the players are organised into teams and required to vote the moves their team will make.
+For this reason, we have decided to take a simple, well known game and develop it into a distributed system: Tic-Tac-Toe played as a peer-to-peer, team-based gaming system, where the players are organised into teams and required to vote the moves their team will make.
 
@@
 
-Upon initialising the game, the player will take a note of their address and port number.
-As the game is played on a local network, the player will attempt to connect to all other existing nodes on the network.
-If none are found, then the client will initialise a new game instance and set itself as one of the leaders.
+A new game instance tries to join an already existing game specified
+by the server name and an RMI Registry handle. If a connection
+succeeds (ie. that machine is paricipating in an active game session),
+the new player will retrieve the current game state, initialise
+differently depending on whether it is assigned to be a leader or not, print the current board state, and prepare for the turn.
 
-If a player manages to connect to another machine on the local network (ie.
-that machine is paricipating in an active game session), they will retrieve the current game state, check if they are one of the leaders, print the current board state, and prepare for the turn.
+The leaders update a common game state at the end of each event before
+broadcasting this updated game state to all team members. Each player
+checks if the received play ends the game. If so, it stops all
+activity and, after the RMI connections close, exits.
 
-The leaders update a common game state at the end of each event before broadcasting this updated game state to all team members.
-Players are told by the leaders when the game, and therefore the session, ends.
+\section{Implementation} % TODO: define RMI
 
-\section{Implementation}
+The implementation follows the RMI distributed object model. A
+\texttt{GameInt} interface extends \texttt{java.rmi.Remote} and lists
+all methods that can be called using RMI.
 
-The implementation follows the RMI distributed object model. A GameInt interface extends java.rmi.Remote and lists remote classes.
+A \texttt{Game} class implements \texttt{GameInt} and extends
+\texttt{UnicastRemoteObject}. It is the main class for each player,
+holding most of its knowledge and behaviour. It provides an
+implementation for all the listed methods in the interface, which
+include the functionality of requesting a player to make a play,
+setting the leader of the game, several methods to access
+role-specific functionality, and a main method that connects to other game instances.
 
-A Game class implements GameInt and extends the UnicastRemoteObject.
-It is an analogy for what a player knows and does during a game.
-It provides an implementation for all the listed classes in the interface which include the functionality of requesting the player to make a play, setting the leader of the game, several methods to interact with the Role of the player, and a main method that connects to other game instances.
+Currently, when running a game, the user supplies a local instance
+name and the name of another instance as command line arguments. Each
+instance then uses this information to do an RMI Registry lookup and
+get a reference to the other peer in an attempt to establish a
+session. This allows multiple different sessions to run simultaneously
+on the same network without interference --- ideal for educational use, where multiple groups are likely to be working on the same internal network.
 
-Currently, when running the game, the user supplies a local instance name and the name of another instance as command line arguments.
-Each instance then uses this information to do an RMI Registry lookup and get a reference to the other client.
-
-The Role interface allows the application to differentiate between Leader and Player classes which extend it.
+% TODO: reference for state design pattern
+The \texttt{Role} interface allows the application to differentiate between \texttt{Leader} and \texttt{Player} classes which extend it.
 It utilises the State design pattern to allow players to seamlessly have their role changed when required (eg.
@@
 
-A Leader is responsible for collecting the votes from all of the members in their team in the turnStarts() method.
-This is done asynchronously using a separate thread for each member – an instance of VoteThread.
-Each thread collects a vote from one assigned player then writes it into a corresponding position in a shared AtomicIntegerArray that they have all been passed by the leader.
-This ensures the main leader thread can then read these values.
+A Leader is responsible for collecting votes from all of the members
+in their team in the \texttt{turnStarts()} method. This is done
+asynchronously using a separate thread for each member --- an instance
+of \texttt{VoteThread}. Each thread collects a vote from one assigned
+player and writes it into a corresponding position in a shared
+\texttt{AtomicIntegerArray}. The leader then uses these votes to decide
+on a play.
 
-Each Leader then informs all players in the game about the most popular vote chosen by their team.
-In the event of two or more moves receiving an equal number of votes, a leader will randomly select  A Player instance represents a regular player that can only vote, it cannot independently generate and broadcast a new game state.
+% TODO: stick to uppercase or lowercase leaders, players
+Each \texttt{Leader} then informs all players in the game about the
+most popular vote chosen by their team. In the event of two or more
+moves receiving an equal number of votes, a leader will randomly
+select one of them. A \texttt{Player} instance represents a regular
+player that can only vote, it cannot independently generate and
+broadcast a new game state.
 
@@
 
-Elections use the bully algorithm to determine which Player instance will become the new Leader instance for their team.
+% TODO: reference the bully algorithm
+Elections use the ``Bully'' algorithm to determine which \texttt{Player} instance will become the new \texttt{Leader} instance for their team.
 The other team members and the leader of the opposing team are all informed of the new leader, creating an altered topology to compensate for the failure:
@@
 	\includegraphics[width=\linewidth]{images/DAS-topology-election.png}
-	\caption{Proposed Topology for UDP-based Version of System}
+	\caption{Proposed Topology for UDP-based Version of System} % TODO: is this caption out of date?
 	\label{fig:topology-election}
@@
 
-If the failed machine joins the game session again, it will not automatically regain Leader status, instead participating as a regular Player instance.
+If the failed machine joins the game session again, it will not automatically regain \texttt{Leader} status, instead participating as a regular \texttt{Player} instance.
 
-All plays are recorded on a GameState object instance which acts as a game board.
-After each turn, GameState checks if the game is over and relays this message to the leaders which can then inform the members of their team.
+All plays are recorded in a \texttt{GameState} object instance, which
+acts as a game board. After each turn, \texttt{GameState} checks if
+the game is over. If it is, the client quietly waits for the RMI
+connections to time out.
 
@@
 Examples of the former generally deal with the difficulties of distributed systems in a bad fashion, promoting questionable design and programming practices.
-One specific example of this is a YouTube guide to building a single player-per-team distributed tic-tac-toe game in Java \cite{MrWayFarOutYoutube15} - while the material is accessible for inexperienced networks programmers, it puts all functionality in a single class and does not comply with any design guidelines.
-As such, this resource has the potential to further confuse learners about programming distributed systems when they try to read more complex documentation which does obey software engineering design principles.
+One specific example of this is a YouTube guide to building a single player-per-team distributed tic-tac-toe game in Java \cite{MrWayFarOutYoutube15} --- while the material is accessible for inexperienced networks programmers, it puts all functionality in a single class and does not comply with any design guidelines.
+As such, this resource has the potential to further confuse learners about programming distributed systems when they try to read more complex documentation, which does obey software engineering design principles.
 
-The latter category consists of resources which provide details of state-of-the-art practices and advanced optimisations.
+The latter category consists of resources that provide details of state-of-the-art practices and advanced optimisations.
 An example of this is a version of the game Quake II, modified to run multiplayer sessions more efficiently over a distributed system \cite{DBLP:conf/nsdi/BharambePS06}.
@@
 The produced software serves as a good demonstration of a distributed game.
-The use of a simple, well known game turned into a distributed, multiplayer experience highlights the benefits and complexities associated with distributed systems, and allows the technology to be demonstrated in an accessible format, allowing users of all levels to make use of the software without being discouraged by an overwhelming level of complexity.
+The use of a simple, well-known game turned into a distributed, multiplayer experience highlights the benefits and complexities associated with distributed systems, and allows the technology to be demonstrated in an accessible format, allowing users of all levels to make use of the software without being discouraged by an overwhelming level of complexity.
 
@@
 
-In the project's current state, leaders are appointed using the bully algorithm when one machine acting as a leader fails.
-However, there is no support for failure of communication from both leaders, i.e.
-if both leaders were to crash at the same time.
-Although the likelihood of this occurring is slim, this possibility still needs to be addressed to improve player experience and to improve the standard of our product as an educational resource.
+In the project's current state, leaders are appointed using the
+``Bully'' algorithm when a leader fails. However, there is no support
+for failure of communication from both leaders, i.e. if both leaders
+were to crash at the same time. Although the likelihood of this
+occurring is slim, this possibility still needs to be addressed to
+improve player experience and to improve the standard of our product
+as an educational resource.
 
-Another improvement which we would make in further iterations would be some form of automatic team balancing - currently, no replacement team members are allocated to an underpopulated team if several peers drop out of a session.
+Another improvement, which we would make in further iterations, would be some form of automatic team balancing --- currently, no replacement team members are allocated to an underpopulated team if several peers drop out of a session.
 Another situation which needs to be addressed is the scenario of all members of a team leaving a game before it is complete, an event which currently has no resolution.

### 2017-12-01 paper/paper.tex
Added one more reference

@@
 	\includegraphics[width=\linewidth]{images/DAS-topology.png}
-	\caption{Topology for Distributed Tic Tac Toe}
+	\caption{Topology for Distributed Tic-Tac-Toe}
 	\label{fig:topology1}
@@
 % TODO: reference for state design pattern
-The \texttt{Role} interface allows the application to differentiate between \texttt{Leader} and \texttt{Player} classes which extend it.
-It utilises the State design pattern to allow players to seamlessly have their role changed when required (eg.
-if the current leader disconnects).
+The \texttt{Role} interface allows the application to differentiate
+between \texttt{Leader} and \texttt{Player} classes which extend
+it. It utilises the State design pattern
+\cite{DBLP:books/daglib/0078642} to allow players to seamlessly have
+their role changed when required (eg. if the current leader
+disconnects).
 
@@
 	\includegraphics[width=\linewidth]{images/DAS-topology-election.png}
-	\caption{Proposed Topology for UDP-based Version of System} % TODO: is this caption out of date?
+	\caption{Topology changes after a leader disconnects}
 	\label{fig:topology-election}

### 2017-12-01 paper/paper.tex
completed performance testing and write-up

@@
 \section{Performance Evaluation}
+We created a script, \texttt{createpeers.py}, to automate the creation of peers for testing system performance.
+This can create a base instance to start the game, and can then create any number of peers which will attempt to connect to the base instance.
+
+We decided to investigate how the ``Bully'' election algorithm scales with the number of peers on a team (fig. 3).
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{images/election-chart.png}
+	\caption{Average time taken to respond to process a response from each peer in an election call}
+	\label{fig:election-img}
+\end{figure}
+
+The general trend is that the election caller will take a longer time, on average, to process each election vote from each peer on its team.
+However, the results for 15 and 20 nodes are anomalous.
+There are several potential causes for this: another process on the machine running these tests blocking one of the voting threads, election threads blocking other threads from accessing a shared resource, and refused connections.
+
+A considerable number of nodes failed during these elections, with this number varying between 4 failures in the 15 node run and 11 in the 30 node run (fig. 4).
+
+\begin{figure}[h]
+	\includegraphics[width=\linewidth]{images/failures.png}
+	\caption{Number of node failures caused by java.rmi.ConnectException}
+	\label{fig:failure-img}
+\end{figure}
+
+These failures were all caused by \texttt{java.rmi.ConnectException}, which was thrown in response to refused TCP connections.
+Java RMI calls require an exclusive TCP connection to a client, which is not suited to the ``Bully'' algorithm.
+Although a new leader is still elected from the suriving nodes, these failures strongly suggest that modifying RMI to use UDP sockets or \texttt{SO\_REUSEADDR}-enabled TCP sockets is required for a more failure-resistant system.
+This idea is covered in more detail below in Architecture Evaluation.
+
+The next step would be to repeat these tests in a scenario closer to how this project is likely to be used in an educational setting such as a university lab exercise: several different machines each running between one and three instances of the game.
+We would record how many failures are caused by \texttt{java.rmi.ConnectException} in comparison to running all instances on \texttt{localhost} as well as comparing the average time taken to process votes.
 
@@
 This requires creating four TCP sockets bound to the same local addr:port combination (ie.
-with SO\_REUSEADDR enabled) to establish a connection with a peer through a proxy server, as described in section 4 of a paper by Ford, et al \cite{ford_srisuresh_kegel_2005}.
+with \texttt{SO\_REUSEADDR} enabled) to establish a connection with a peer through a proxy server, as described in section 4 of a paper by Ford, et al \cite{ford_srisuresh_kegel_2005}.
 Not only is this a significant overhead, with four sockets required for each peer a client wants to make contact with, it is also more complex to perform than the UDP equivalent - our initial tests could not establish contact between peers using the instructions described in the Ford paper.

### 2017-12-01 createpeers.py

@@ -6,7 +6,11 @@ import threading
 
 def client_thread(i, move):
     #Create client, execute move on repeat
-    subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost; sleep 5; while true; do %d; done\"" % (i, move), shell=True)
+    subprocess.call("gnome-terminal -x sh -c \"yes %d | java Game client%d server localhost\"" % (i, move), shell=True)
+    
+def client_thread_no_move(i):
+    #Create client, execute move on repeat
+    subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost\"" % i, shell=True)
     
 
 if sys.argv[1] == "server":
@@ -29,10 +33,27 @@ elif sys.argv[1] == "clients":
             c_thread.start()
             #subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost\"" % i, shell=True)
             #sleep to allow time for server to resolve connections
-            time.sleep(5)
+            time.sleep(1)
+        print("%d peers created" % peer_count)
+    except Exception as e:
+        print(e)
+        
+elif sys.argv[1] == "clients_no_move":
+    try:
+        #number of peers to create in the session
+        peer_count = int(sys.argv[2])
+        for i in range(peer_count):
+            #create instance threads named client0, client1, etc. in a new window
+            print("Creating peer thread %d" % i)
+            move = random.randint(0,8)
+            c_thread = threading.Thread(target=client_thread_no_move, args=(i,))
+            c_thread.start()
+            #subprocess.call("gnome-terminal -x sh -c \"java Game client%d server localhost\"" % i, shell=True)
+            #sleep to allow time for server to resolve connections
+            time.sleep(1)
         print("%d peers created" % peer_count)
     except Exception as e:
         print(e)
 
 else:
-    print("Usage:\npython server - creates base server instance on localhost\npython clients [number] creates [number] of clients to connect to server instance")
+    print("Usage:\npython server - creates base server instance on localhost\npython clients [number] - creates [number] of clients to connect to server instance\npython clients_no_move [number] - creates [number] of clients to connect to server instance without making any moves")

### 2017-12-01 paper/paper.tex
added texttt labels to a few places

@@
 Java RMI calls require an exclusive TCP connection to a client, which is not suited to the ``Bully'' algorithm.
-Although a new leader is still elected from the suriving nodes, these failures strongly suggest that modifying RMI to use UDP sockets or \texttt{SO\_REUSEADDR}-enabled TCP sockets is required for a more failure-resistant system.
+Although the election algorithm itself is resilient and still elects a new leader from the surviving nodes, these failures strongly suggest that modifying RMI to use UDP sockets or \texttt{SO\_REUSEADDR}-enabled TCP sockets is required for a more failure-resistant peer-to-peer system.
 This idea is covered in more detail below in Architecture Evaluation.
@@
 The next step would be to repeat these tests in a scenario closer to how this project is likely to be used in an educational setting such as a university lab exercise: several different machines each running between one and three instances of the game.
-We would record how many failures are caused by \texttt{java.rmi.ConnectException} in comparison to running all instances on \texttt{localhost} as well as comparing the average time taken to process votes.
+We would record how many failures are caused by \texttt{java.rmi.ConnectException} in comparison to running all instances on \texttt{localhost} as well as comparing any difference in average time taken to process votes.
 
@@
 
-Hole punching for NAT traversal requires the use of an external, publicly-accessible proxy server to distribute the public-facing IP address and port combinations (addr:port) of NAT-obscured clients which want to establish a peer-to-peer session.
+Hole punching for NAT traversal requires the use of an external, publicly-accessible proxy server to distribute the public-facing IP address and port combinations (\texttt{addr:port}) of NAT-obscured clients which want to establish a peer-to-peer session.
 This traversal method is used by many popular types of distributed applications such as multiplayer gaming, VoIP, and peer-to-peer file sharing.
@@
 As mentioned above, Java RMI uses TCP for its operations by default.
-While it is possible to modify the sockets created by RMI calls to use UDP instead, this requires extensive work using RMISocketFactory which is beyond the time allowed by the project.
+While it is possible to modify the sockets created by RMI calls to use UDP instead, this requires extensive work using \texttt{RMISocketFactory} which is beyond the time allowed by the project.
 
